### 2016-12-17 Latex/article.tex
First commit

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfm}
%\documentclass[referee]{jfm} %for double spaced output for submission

\usepackage{graphicx}
\usepackage{natbib}

\usepackage{color}
\usepackage{soul}

\pdfoptionpdfminorversion 4

% \pdfcompresslevel0

% See if the author has AMS Euler fonts installed: If they have, attempt2
% to use the 'upmath' package to provide upright math.
\ifCUPmtlplainloaded \else
  \checkfont{eurm10}
  \iffontfound
    \IfFileExists{upmath.sty}
      {\typeout{^^JFound AMS Euler Roman fonts on the system,
                   using the 'upmath' package.^^J}%
       \usepackage{upmath}}
      {\typeout{^^JFound AMS Euler Roman fonts on the system, but you
                   dont seem to have the}%
       \typeout{'upmath' package installed. JFM.cls can take advantage
                 of these fonts,^^Jif you use 'upmath' package.^^J}%
       \providecommand\upi{\pi}%
      }
  \else
    \providecommand\upi{\pi}%
  \fi
\fi

% See if the author has AMS symbol fonts installed: If they have, attempt
% to use the 'amssymb' package to provide the AMS symbol characters.

\ifCUPmtlplainloaded \else
  \checkfont{msam10}
  \iffontfound
    \IfFileExists{amssymb.sty}
      {\typeout{^^JFound AMS Symbol fonts on the system, using the
                'amssymb' package.^^J}%
       \usepackage{amssymb}%
       \let\le=\leqslant  \let\leq=\leqslant
       \let\ge=\geqslant  \let\geq=\geqslant
      }{}
  \fi
\fi

% See if the author has the AMS 'amsbsy' package installed: If they have,
% use it to provide better bold math support (with \boldsymbol).

\ifCUPmtlplainloaded \else
  \IfFileExists{amsbsy.sty}
    {\typeout{^^JFound the 'amsbsy' package on the system, using it.^^J}%
     \usepackage{amsbsy}}
    {\providecommand\boldsymbol[1]{\mbox{\boldmath $##1$}}}
\fi

%%% Example macros (some are not used in this sample file) %%%

% For units of measure
\newcommand\dynpercm{\nobreak\mbox{$\;$dyn\,cm$^{-1}$}}
\newcommand\cmpermin{\nobreak\mbox{$\;$cm\,min$^{-1}$}}

% Various bold symbols
\providecommand\bnabla{\boldsymbol{\nabla}}
\providecommand\bcdot{\boldsymbol{\cdot}}
\newcommand\biS{\boldsymbol{S}}
\newcommand\etb{\boldsymbol{\eta}}

% For multiletter symbols
\newcommand\Real{\mbox{Re}} % cf plain TeX's \Re and Reynolds number
\newcommand\Imag{\mbox{Im}} % cf plain TeX's \Im
\newcommand\Rey{\mbox{\textit{Re}}}  % Reynolds number
\newcommand\Pran{\mbox{\textit{Pr}}} % Prandtl number, cf TeX's \Pr product
\newcommand\Pen{\mbox{\textit{Pe}}}  % Peclet number
\newcommand\Ai{\mbox{Ai}}            % Airy function
\newcommand\Bi{\mbox{Bi}}            % Airy function

% For sans serif characters:
% The following macros are setup in JFM.cls for sans-serif fonts in text
% and math.  If you use these macros in your article, the required fonts
% will be substitued when you article is typeset by the typesetter.
%
% \textsfi, \mathsfi   : sans-serif slanted
% \textsfb, \mathsfb   : sans-serif bold
% \textsfbi, \mathsfbi : sans-serif bold slanted (doesnt exist in CM fonts)
%
% For san-serif roman use \textsf and \mathsf as normal.
%
\newcommand\ssC{\mathsf{C}}    % for sans serif C
\newcommand\sfsP{\mathsfi{P}}  % for sans serif sloping P
\newcommand\slsQ{\mathsfbi{Q}} % for sans serif bold-sloping Q

% Hat position
\newcommand\hatp{\skew3\hat{p}}      % p with hat
\newcommand\hatR{\skew3\hat{R}}      % R with hat
\newcommand\hatRR{\skew3\hat{\hatR}} % R with 2 hats
\newcommand\doubletildesigma{\skew2\tilde{\skew2\tilde{\Sigma}}}
%       italic Sigma with double tilde

% array strut to make delimiters come out right size both ends
\newsavebox{\astrutbox}
\sbox{\astrutbox}{\rule[-5pt]{0pt}{20pt}}
\newcommand{\astrut}{\usebox{\astrutbox}}

\newcommand\GaPQ{\ensuremath{G_a(P,Q)}}
\newcommand\GsPQ{\ensuremath{G_s(P,Q)}}
\newcommand\p{\ensuremath{\partial}}
\newcommand\tti{\ensuremath{\rightarrow\infty}}
\newcommand\kgd{\ensuremath{k\gamma d}}
\newcommand\shalf{\ensuremath{{\scriptstyle\frac{1}{2}}}}
\newcommand\sh{\ensuremath{^{\shalf}}}
\newcommand\smh{\ensuremath{^{-\shalf}}}
\newcommand\squart{\ensuremath{{\textstyle\frac{1}{4}}}}
\newcommand\thalf{\ensuremath{{\textstyle\frac{1}{2}}}}
\newcommand\Gat{\ensuremath{\widetilde{G_a}}}
\newcommand\ttz{\ensuremath{\rightarrow 0}}
\newcommand\ndq{\ensuremath{\frac{\mbox{$\partial$}}{\mbox{$\partial$} n_q}}}
\newcommand\sumjm{\ensuremath{\sum_{j=1}^{M}}}
\newcommand\pvi{\ensuremath{\int_0^{\infty}%
  \mskip \ifCUPmtlplainloaded -30mu\else -33mu\fi -\quad}}

\newcommand\etal{\mbox{\textit{et al.}}}
\newcommand\etc{etc.\ }
\newcommand\eg{e.g.\ }

\newcommand{\uu}{\textbf{u}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\JJ}{\textbf{J}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ff}{\textbf{f}}
\newcommand{\FF}{\textbf{F}}
\newcommand{\baa}{\textbf{a}}
\newcommand{\bb}{\textbf{b}}
\newcommand{\NN}{\textbf{N}}

\newcommand{\eps}{\varepsilon}

\newcommand{\ErtelPV}{\mathcal{Q}}

\newcommand{\fdiss}{f_{\mbox{\scriptsize diss}}}

\newcommand{\D}{\mbox{D}}

\newcommand{\eez}{\boldsymbol{e_z}}
\newcommand{\scalarprod}[2]{\big( #1 \, , \ #2 \big)_{\kk}}

\newcommand{\mean}[1]{\langle #1 \rangle}

\newcommand{\meane}[1]{\langle #1 \rangle}
% \newcommand{\meane}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}e}}}

\newcommand{\meanx}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\mbox{\scriptsize$\xx$}}}}

\newcommand{\meant}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\theta}}}

\newcommand{\means}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}s}}}

\newcommand{\shocks}{ {\mbox{\tiny shocks}} }

\newcommand{\kmax}{k_{\mbox{\scriptsize max}}}
\newcommand{\kdiss}{k_{\mbox{\scriptsize diss}}}

\newcommand{\mA}{\mathcal{A}}

\newlength{\halfwidth}
\setlength{\halfwidth}{66mm}

\newcommand{\PA}[1]{{\color{green}#1}}

\newcommand{\Add}[1]{{\color{blue}#1}}
% \newcommand{\Add}[1]{#1}
\newcommand{\Remove}[1]{{\color{red}\st{#1}}}
% \newcommand{\Remove}[1]{}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title[Wave energy cascade in shallow-water flows]%
{Wave energy cascade in forced-dissipative one-layer shallow-water flows}

\author[P. Augier and E. Lindborg]%
{Pierre Augier$^{1,2}$%
  \thanks{Email address for correspondence: p.augier@damtp.cam.ac.uk},\ns
and Erik Lindborg$^1$}

\affiliation{$^1$Linn\'e Flow Centre, KTH Mechanics,
SE-100 44 Stockholm, Sweden\\[\affilskip]
$^2$Department of Applied Mathematics and Theoretical Physics,
University of Cambridge, Cambridge CB3 0WA, United Kingdom}

\pubyear{2013}
\volume{???}
\pagerange{???--???}
% Do not enter received and revised dates. These will be entered by
% the editorial office.
\date{?; revised ?; accepted ?. - To be entered by editorial office}
%\setcounter{page}{1}
\begin{document}

\maketitle

% \baselineskip 8mm
\begin{abstract}
The dynamics of one-layer shallow-water flows forced in medium-scale
wave modes and dissipation acting at small scales is investigated by
means of theoretical analysis and numerical simulations with
resolutions up to $7680^2$.
%
We first focus on the non-rotating case for which a statistically
stationary regime is obtained for all resolutions and Froude numbers
$F_f = \eps/(k_f c)$, where $\eps$ is the energy dissipation rate,
$k_f$ the forcing wave number and $c$ the wave speed.
%
All the injected energy goes into a downscale energy
cascade. Consequently, the spectral energy flux is constant and
positive between the forcing wave number and the wave numbers where
dissipation is active.  The spectral energy flux is exactly
equipartioned between kinetic and potential energy.  An exact third
order structure function law of Kolmogorov type is derived and
numerically verified.  However, the downscale cascade is less
efficient than in three-dimensional Navier-Stokes turbulence, since
the mean energy of the statistically stationary flow scales as
$\sqrt{\eps c /k_f}$ and varies with the dissipative scale.
%
Moreover, the spectra scale as $k^{-2}$ and the structure functions of
order larger than two scale approximately as $r$.
%
%
This intermittent scaling is explained by a simple model based on the
assumption that the statistics is totally determined by shocks.
%
It is verified that the dynamics is not drastically modified by system
rotation as long as the forcing wave number is larger than the
deformation wave number $k_d = f/c$, where $f$ is the Coriolis
parameter.
\end{abstract}

% \begin{keywords}
% \end{keywords}

\section{Introduction}

The shallow-water models have been widely used to study basic
mechanisms occurring in geophysical flows \cite[see for
example][]{VallisLIVRE2006}.
%
These models are based on the hydrostatic approximation which is
extremely well satisfied for a very wide range of scales in the oceans
and in the atmosphere \Add{and on which are based the so-called
primitive equations}.  They also rely on a stronger hypothesis
regarding the density stratification.  The fluid is assumed to be
structured in a limited number of thin homogeneous layers, which leads
to equations involving only two-dimensional operators. \Add{However,
as long as there is no overturning, the shallow-water model in the
limit of a large number of layers and the primitive equations describe
the same dynamics. In the opposite limit, two models are particularly
important from the theoretical point of view.} The two-layer
shallow-water model captures the baroclinic instability which is of
primary importance for the dynamics of the oceans and of the
atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
%
The equations of the one-layer shallow-water model are very similar to
the two-dimensional compressible Navier-Stokes equations.  The
one-layer model does not capture the baroclinic instability but it is
one of the simplest hydrodynamic models with coexisting eigenmodes of
the linearized operator with zero linear frequency and with finite
non-zero linear frequency, $\omega = \pm \omega_l = \pm\sqrt{f^2 + c^2
k^2}$, where $f$ is the Coriolis parameter, $c$ the wave speed and $k
= |\kk|$ the wave number.
%

For geophysical applications, it is convenient to consider the limits
for which the linear frequency $\omega_{l}$ is much larger than the
characteristic nonlinear frequency $\omega_{nl} = U k$, where $U$ is a
characteristic velocity.  In the limit $f/\omega_{nl} \sim
ck/\omega_{nl} \rightarrow \infty$, one can obtain a quasi-geostrophic
one-layer shallow-water model, i.e. derive self-consistent evolution
equations for the 0-linear-frequency variable, the Charney potential
vorticity.  In this limit, the dynamics of the 0-linear-frequency
variable is totally decorrelated from the fast waves.
%
For finite rotation and wave speed, there are slow balanced motions,
which are associated with relatively weak non-zero linear frequency
modes.
%
When $\omega_{nl}(k) \ll \omega_{l}(k) $, there is only a weak
interaction between waves and the balanced modes.
%
Due to its relative simplicity, the one-layer shallow-water model has
been used to study issues like the projection on the balanced manifold
\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
waves by balanced flows \cite[]{FargeSadourny1989,
LahayeZeitlin2012,Vanneste2013}. \Add{In this study, we consider the
same model in order to address a related but different issue.}

Balanced flows have a dynamics which is very similar to
two-dimensional turbulence.
%
In the quasi-geostrophic limit, the space-averaged energy and the
space-averaged potential enstrophy are conserved, positive definite
and proportional in Fourier space. This leads\Add{, as in
two-dimensional turbulence,} to an upscale energy cascade and a
downscale enstrophy cascade.
%
Thus, the small-scale dissipation vanishes in the limit of zero
viscosity.
%
In contrast to this, statistical mechanics indicates that wave energy
should be transferred towards small scales \cite[]{Warn1986}.
Numerical simulations have shown that long waves lose their energy by
transferring it to shorter waves \cite[]{Sadourny1975,
FargeSadourny1989, YuanHamilton1994}.  The mechanism of a wave cascade
has been proposed to explain the attraction towards the balanced flows
\cite[]{Sadourny1975}.  \cite{YuanHamilton1994} showed that
statistically stationary shallow-water flows can be obtained forcing
quasi-geostrophic modes at large scales and with dissipation only at
small scales.
%
\Add{They propose that the $k^{-5/3}$ spectra that are observed at the
atmospheric meso-scales} \cite[]{NastromGage1985, FrehlichSharman2010}
\Add{and reproduced by some atmospheric models based on the primitive
equations} \cite[]{KoshykHamilton2001, Skamarock2004,
Hamilton_etal2008} \Add{could be reproduced by the one-layer
shallow-water model by a downscale energy cascade of waves.}

However, few studies have investigated the dynamics of the downscale
energy cascade of waves in shallow-water models.
%
Most numerical studies have investigated decaying flows and have not
focused on the cascade \cite[]{FargeSadourny1989,
LarichevMcWilliams1991, SpallMcWilliams1992,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
%
Moreover, there are strong discrepancies between different numerical
results and between these results and available theoretical results.
%
\cite{FargeSadourny1989} reported spectra scaling like $k^{-4}$
whereas those shown by \cite{LahayeZeitlin2012} scale like $k^{-6}$.

Some predictions are based on the weak wave turbulence formalism
\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011}.  In the case of
inertial-gravity waves, \cite{FalkovichMedvedev1992} showed that the
exact solutions of the kinetic equation corresponding to a constant
downscale energy flux is associated with a spectrum scaling like $E(k)
\sim P^{2/3} k^{-8/3}$.
%
However, as pointed out by \cite{FalkovichMeyer1996} when the
dispersion is small enough, waves moving in the same direction have
the same velocity and interact strongly, which can lead to the
formation of shocks. \Add{In this case, the weak wave turbulence
formalism is unable to give valuable results.}  Such sharp and long
structures have been observed in many simulations
\cite[]{FargeSadourny1989, PolvaniMcWilliamsSpallFord1994,
LahayeZeitlin2012} but \Add{their effects} are not well understood.
%
\cite{Kuznetsov2004} predicted that a flow dominated by shocks should
be associated with spectra scaling like $E(k) \sim k^{-2}$.
%
\cite{FalkovichMeyer1996} simulated acoustic turbulence dominated by shocks 
and reported spectra in agreement with this theoretical prediction.
%
The shocks also lead to very strong intermittency, as demonstrated by
\cite{BouchaudMezardParisi1995} and \cite{WeinanKhaninMazelSinai1997}
in the case of Burgers turbulence.
%

The dynamics of waves in complex shallow-water flows is not well
understood.  \Add{Our long-term goal is to understand this dynamics
which involves interactions between balance and unbalance parts of the
flow. However, even the wave cascade without balance flow is not well
understood.} A statistically stationary wave cascade has never been
simulated at high resolution.
%
Is it possible to obtain a statistically stationary state in the very
fast wave regime, and how does the mean energy in this hypothetical
regime depends on the forcing and the wave speed?  
% What is the slope of 
\Add{How does} the energy spectra \Add{scale as a function of the
input parameters}?
%
All these questions are open and straightforward simulations may
provide the answers.

\cite{YuanHamilton1994} and \cite{FalkovichMeyer1996} carried out
forced dissipated simulation of the one-layer shallow water model and
acoustic turbulence, respectively.
% 
However, the resolution in both studies was quite coarse \Add{($386^2$
and $101^2$ grid points, respectively)} so that the inertial range was very
narrow.  Moreover, \cite{FalkovichMeyer1996} used a very particular
type of strongly anisotropic forcing \Add{with waves in only one
direction} while \cite{YuanHamilton1994} forced in geostrophic modes
which severely complicates the interpretation.

In this paper, we investigate the dynamics of one-layer shallow-water
flows forced in the wave modes by means of high resolution numerical
simulations \Add{with up the  $7680^2$ grid points}.
%
The model is described in section~\ref{section_basic_theory}.
%
We then derive some new theoretical results for this model and for
turbulence dominated by shocks
(section~\ref{section_advanced_theory}).
%
The numerical methods and the forcing are described in
section~\ref{section_methods}.
%
The main numerical results corresponding to the simple case without
global rotation are presented in
section~\ref{section_wave_cascade_f=0}.
%
We investigate the sensibility of the non-rotating results to a
moderate rotation (section~\ref{section_rotation}).

\section{The one-layer shallow-water model}
\label{section_basic_theory}
% \input{section_basic_theory}

\section{Theoretical analysis}
\label{section_advanced_theory}
% \input{section_advanced_theory}

\section{Numerical methods}
\label{section_methods}
% \input{section_methods}

\section{Wave cascade without global rotation}
\label{section_wave_cascade_f=0}
% \input{section_wave_cascade_f=0}

\section{Effects of global rotation}
\label{section_rotation}
% \input{section_rotation}

\section{Conclusions}

We have numerically investigated one-layer shallow-water flows forced
with medium-scale waves and dissipated at small scales.
%
First, we have focused on the non-rotating case and varied the
resolution and the wave speed over a wide range.
%
The spectral energy budget for the full non-quadratic energy has been
derived and has allowed us to show that the KE and the APE fluxes are
constant and equal over a wide range of wave numbers.
%
The KE and APE spatio-temporal spectra present equal and very strong
peaks at the linear wave frequency.
%
These two results show that there is a downscale energy cascade of
non-dispersive waves, explaining the generation of statistically
stationary flows.
%
We have derived an exact Kolmogorov law for the one-layer
shallow-water flow based on the assumption that there is a constant
energy flux due to a purely divergent flow and verified that the
numerical results are in agreement with this theoretical law.

The time- and spaced-averaged energy scales like $E = C_n \sqrt{\eps
L_f c}$ and increases with the resolution at least up to $n = 5760$.
The spectra scale like the Zakharov-Sagdeev spectrum $k^{-3/2}
\sqrt{\eps c}$ at relatively small wave numbers $k \lesssim 2 k_f$ and
like $k^{-2} c^{1/3} \eps^{5/9} k_f^{4/9}$ over the inertial range.
We stress that we lack any convincing theoretical explanations for
these scaling laws and we hope our results will motivate theoretical
investigations on these challenging issues.

Nevertheless, the $k^{-2}$ scaling can be explained by the presence of
discontinuities \cite[]{Kuznetsov2004}.  The snapshots indeed show
very sharp hydraulic jumps, which are the equivalent of shocks in
compressible flows.
%
Following \cite{BouchaudMezardParisi1995} and
\cite{WeinanKhaninMazelSinai1997}, we have built a simple model based
on the assumption that the structure functions are determined only by
shocks.
%
Using only this hypothesis, we explain the extreme intermittent
characteristics of the flows, i.e.\ the $r^1$-scaling of the structure
functions of order larger than 2 and the $r^{-1}$-scaling of the
flatness.
%
Adding a simple hypothesis on the structure of the velocity field
across the shocks, namely $\delta u_L = \delta u \sin\theta$ and
$\delta u_T = \delta u \cos\theta$, where $\theta$ is the angle
between the shock line and the separation vector, we have predicted
the values of ratios of structure functions $R_p(r) \equiv
\mean{|\delta u_L|^p}/\mean{|\delta u_T|^p}$ and of the flatness
factors $F_T/F_L = 1.5$.
%
% Very interestingly, the flatness factors computed from atmospheric
% data \cite[]{Lindborg1999} are also very large with a slope close to
% $-1$ and a ratio $F_T/F_L$ equal to $1.5 \pm 0.1$.
%
Finally, we have shown that as long as $k_f>k_d$, the dispersion added
by the rotation does not prevent the production of the regime studied
in the non-rotating case and only weakly modify the results on the
wave cascade.

% In order to draw conclusions on the interpretation of turbulent
% measurements in geophysical flows, we first have to stress that the
% one-layer shallow-water model can only give insights on some
% mechanisms important for geophysical fluids but is not suitable for
% simulate the oceans and atmosphere.
% %
% Most of the simulations of theses systems are performed solving the
% so-called primitive equations, which are (or in the atmospheric case
% are very similar to) the Navier-Stokes equations under the Boussinesq
% and the hydrostatic approximations.
% %
% We focus on the non-rotating case since it is a reasonable
% approximation for the medium scales for which an energy cascade seems
% to strongly influence the dynamics of geophysical flows, namely the
% mesoscales in the atmosphere and the sub-mesoscales in the oceans.
% %
% The dominant balance for the primitive equations in the strongly
% stratified and high Reynolds number limits corresponds to flows
% structured in thin layers \cite[]{BillantChomaz2001}.
% %
% This layering, which has been widely observed in nature, is of course
% inconsistent with the one-layer shallow-water model.
% %
% Another aspect of the dominant balance in the strongly stratified and
% high Reynolds number case is that the energetic waves are not fast
% compared to the balanced flows ($\omega_l \sim \omega_{nl}$), which is
% very different from the case of the studied one-layer shallow-water
% flows.
% %
% Moreover, the strongly stratified and high Reynolds number dominant
% balance would corresponds in terms of the shallow-water Froude number
% to $Fr \sim 1$.  This limit can not be correctly simulated even with a
% multi-layer shallow-water model since when $Fr \sim 1$ the thickness
% of the layers tends to reach the value 0 for which the shallow-water
% models cease to be well defined.
% %
% Therefore, the shallow-water models are not suitable for simulating
% the geophysical medium scales where the cascade is important.

Our results on the dynamics of the wave energy cascade in the
one-layer shallow-water model can yet give interesting informations on
the interpretation of turbulent measurements in geophysical fluids.
%
Since we have shown in this case that many statistical results are
dominated by the effects of the discontinuities due to shocks, we can
conjecture that the discontinuities or front in the atmosphere can
significantly influence the cascade processes and the statistical
results.
%
It would be interesting in the future to quantity the effects of the
processes of frontogenesis in terms of downscale energy cascade in
geophysical flows.

\section*{Acknowledgements}

The authors thank Gregory Falkovich and Sergey Nazarenko for valuable
comments and suggestions regarding the interpretation of the
simulations.

\appendix

\section{Kinetic energy dissipation and forcing rates}
\label{app_comp}
% \input{appendix_dissip_forcing}

\bibliographystyle{jfm}
\bibliography{biblio}

\end{document}