### 2018-10-16 Latex/section_methods.tex

We have carried out a series of twenty forced-dissipative simulations of the SW equations in a quadratic domain with periodic boundary conditions using a standard pseudo spectral code. The simulations were started from zero initial conditions and were run till a statistically stationary state was reached. The equations 
were reformulated in terms of  the three eigenmodes of the linearized equations in Fourier space. One of these eigenmodes is zero for an irrotational flow.  
We first carried out a few test simulations in which all three eigenmodes were allowed to evolve in time. However, when only the two wave modes are forced vorticity remains very small throughout the simulations and does not affect the wave dynamics. Therefore, we chose to solve the equations only for the two wave modes. As discussed in the previous section,  a linear viscous term in (\ref{eq_uu}) 
will not conserve momentum exactly and does not give a strictly positive definite expression for energy dissipation. We have relaxed the requirements of exact momentum conservation and positive definite energy dissipation. 
Instead of  using the nonlinear form of the viscous term in (\ref{eq_uu}) we have used a higher order linear operator  $ \nu_8 \nabla^{8} $ both in (\ref{eq_uu}) and
(\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion' or a `hyperviscosity'. There is no strong physical motivation for this
modification but, as discussed by \cite{FargeSadourny1989}, a molecular viscous operator 
 of the form $\nu \nabla^2 $ is not necessarily a
relevant model of the actual small-scale dissipation for SW
flows, which should rather be described as a transition from
two-dimensional to three-dimensional motions before reaching the
scales where dissipation actually occurs.
%
Time advancement is carried out by a classical fourth-order
Runge-Kutta scheme for the nonlinear term and an exact integration for
the linear and dissipative terms.  This explicit integration is
especially useful at very high resolution and very large $c$ when
the shortest waves are very fast. 
We use an adaptable time step method which maximizes the time step
over a standard Courant-Friedrichs-Lewy condition
\cite[]{Lundbladh1999, AugierChomazBillant2012}.
%
Most of the aliasing is removed by truncating 8/9 of the modes along
each direction \cite[for a detail discussion on the issues of the
non-conservation of the non-quadratic energy and the aliasing errors
in the truncated one-layer shallow water model,
see][]{FargeSadourny1989}. The resolution is characterized by the number of nodes, $n$, in the each
direction and has been varied from 960 to 7680.
%
The wave speed $c$ has been varied over almost two orders of magnitude from
10 to 700. The wave modes are forced in a shell in spectral space 
corresponding to relatively small wave numbers, 
$ 5 \delta k <  |\kk| <  8 \delta k$,  
where  $ \delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic domain. 
The forcing is white noice in time and designed so that the energy injection rate is normalized to unity at each time step.
The energy is dissipated at the smallest resolved scales by hyperviscosity.
The value of the
hyperviscosity, $\nu_8$, is chosen in such a way that the  maximum wave number is larger than the dissipation wave number
\begin{equation} 
k_{d} = \left( \frac{{\nu_8}^3}{\eps} \right)^{-\frac{1}{22}},
\end{equation}
 where $ \eps $ is the mean energy dissipation rate. We found that  $
 k_{d} \simeq \kmax/ 2.5$, where $\kmax =
(8/9)\pi n/L $ is the maximum wave number, is  sufficient in order for dissipation spectra to be resolved. Since the energy injection rate is normalized to unity and all simulations reach a stationary state in which dissipation is equal to the injection rate, $ \nu_8 $ can be chosen already from the start so that this condition will be fulfilled in the stationary state.  The strength of the stratification is characterized by a Froude number, defined as
\begin{equation} \label{Fr}
F_f = \frac{\varepsilon^{1/3}}{c k_f^{1/3}} \, ,
\end{equation} 
a parameter that can be determined from the start of each simulations. 
The  simulation parameters are listed in table 1.

\begin{table}
\begin{center}

\label{Table1}
\begin{tabular}{lrrrrrrrr}

\toprule
{} &   $n$ &  $c$ &  $\nu_8$ &  $\eps$ &  $ k_{d}  / {k_f} $ &   $F_f$ &    $t_{\max}$ \\
\midrule
W1  &   960 &   10 & 1.56e-10 &        1.04 &                    28.8 &   0.111 &                49.9 \\
W2  &  1920 &   10 & 9.68e-13 &        1.06 &                    57.7 &   0.112 &               49.8 \\
W3  &  3840 &   10 &    6e-15 &         1.10&                     115 &   0.113 &               49.9 \\
W4  &  7680 &   10 & 3.72e-17 &        1.18 &                     231 &   0.116 &               49.9 \\
W5  &   960 &   20 & 1.56e-10 &           1.00 &                    28.8 &   0.055 &              49.9 \\
W6  &  1920 &   20 & 9.69e-13 &        1.02 &                    57.7 &  0.0553 &                120 \\
W7  &  3840 &   20 & 6.01e-15 &        1.08 &                     115 &  0.0563 &                120 \\
W8  &  7680 &   20 & 3.72e-17 &        1.21 &                     231 &  0.0585 &                 120 \\
W9  &   960 &   40 & 1.56e-10 &          1.00 &                    28.8 &  0.0274 &                49.8 \\
W10 &  1920 &   40 & 9.68e-13 &       1.00 &                    57.7 &  0.0274 &               49.8 \\
W11 &  3840 &   40 &    6e-15 &        1.06 &                     115 &   0.028 &               49.8 \\
W12 &   960 &  100 & 1.56e-10 &       1.00 &                    28.8 &   0.011 &                49.8 \\
W13 &  1920 &  100 & 9.69e-13 &        1.01 &                    57.7 &   0.011 &                 120 \\
W14 &  3840 &  100 & 6.01e-15 &        1.06 &                     115 &  0.0112 &                 120 \\
W15 &  7680 &  100 & 3.72e-17 &        1.22 &                     231 &  0.0117 &                95.8 \\
W16 &   960 &  400 & 1.56e-10 &        1.14 &                    28.8 & 0.00287 &                 120 \\
W17 &  1920 &  400 & 9.69e-13 &        1.15 &                    57.7 & 0.00288 &                 120 \\
W18 &  3840 &  400 & 6.01e-15 &        1.12 &                     115 & 0.00286 &                119 \\
W19 &   960 &  700 & 1.56e-10 &        1.27 &                    28.8 &  0.0017 &                49.8 \\
W20 &  1920 &  700 & 9.68e-13 &        1.05 &                    57.7 &  0.0016 &                49.8 \\
\bottomrule

\end{tabular}
\caption{Parameters for all simulations. $ n $: number of nodes in each direction, $ c $: wave speed, $ \nu_8 $: hyperviscosity, $ \eps $: time averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio between dissipation wave number and forcing wave number, $ F_f  $: Froude number, $ t_{max} $:   end time of simulation.}
\end{center}
\end{table}

\begin{figure}
\centerline{\includegraphics[width=5.12in]{../Pyfig/fig_Emean_time}}
\caption{Space averaged energy, 
$E = E_K + E_A $, versus time 
for all simulations. }
\label{fig_Evstime}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=5.8in]{../Pyfig/fig_energy_w}}
\caption{Left: Mean energy in the stationary state versus $ c $ for different resolutions. The lines have slope $ 0.5 $, corresponding to $ E \propto c^{1/2} $. Right: Normalized mean energy in the stationary state versus resolution $ n $ for different $ c $.}
\label{MeanE}
\end{figure}

The time evolution of  
mean energy is shown in
figure~\ref{fig_Evstime} for all twenty simulations. After an initial period of growth, mean energy is leveling out at different values for different runs.  This is not the same picture as we see in simulations of three-dimensional incompressible turbulence or stratified turbulence, where all curves would have leveled out more or less at the same value. We find that mean energy in the stationary state is a function of both $ c $ (or the Froude number) and resolution $ n $ and that the functional dependence  can be approximately written as
\begin{equation} \label{MeanEnergy}
E \propto \sqrt{L_f c \eps} \, n^{\alpha}  \, ,
\end{equation} 
where $ L_f = \pi/k_f $ is the forcing scale. In figure \ref{MeanE} to the left, we see $ E/\sqrt{\eps} $ as function of $ c $ for different resolutions $ n $. For each resolution the  points are approaching a power law $ C_n c^{1/2} $  with increasing values of $ C_n $ for increasing $ n $. In the same figure to the right we see 
$ E/\sqrt{L_f c \eps} $ for different values of $ c $. For each $ c $ the points follow a power law $ n^{\alpha} $, where $ \alpha $ is becoming slightly larger with increasing $ c $, approaching almost the same value $ \alpha \approx 0.33 $ for the three highest values of $ c $.  The relation (\ref{MeanEnergy}) can be reformulated as
\begin{equation} \label{Dissipation}
\frac{\eps} {(E^{3/2} / L_f)} \propto \frac{E^{1/2}}{c} n^{-2\alpha}
\end{equation} 
Since the simulations are well resolved the dependence on resolution should not be interpreted as consequence of the numerics but rather as a Reynolds number effect, although it is not hundred percent clear how a Reynolds number should be defined in our case.  We may note, however, that $ n $ is  proportional $ k_{d} / k_f $ and  the corresponding ratio in Kolmogorov turbulence scales as $ Re^{3/4} $ since the Kolmogorov scale is proportional to $ \nu^{3/4} $, where $ \nu $ is the kinematic viscosity. If we simply define a Reynolds number as
\begin{equation}
Re = \left ( \frac{k_d} {k_f} \right )^{4/3} \, ,
\end{equation} 
we can substitute $ n^{-2\alpha} $ in (\ref{Dissipation}) with $ Re^{-8 \alpha/3} $.  The normalised energy dissipation on the left hand side of  (\ref{Dissipation}) will thus go to zero in the limit of large Reynolds number. This is very different from three-dimensional turbulence 
where there is a local Richardson-Kolmogorov cascade, in which energy is successively transfered from large to small scales of motion.  That the cascade is local means that eddies of a particular scale are not directly influenced by eddies of widely different scales. As a consequence, the normalized dissipation is of the order of unity independent of Reynolds number \cite[]{Pope, TennekesLumley}. A review of the experimental evidence and theoretical implications of this fundamental law of Kolmogorov turbulence is given by \cite{Vassilicos2015}. 
Our result (\ref{Dissipation})  suggests that SW wave turbulence does not fit into the paradigm of a local Richardson-Kolmogorov cascade. Moreover, it is not only in the limit of large Reynolds number the normalized dissipation will go to zero but also in the limit of strong stratification, $ E^{1/2}/c \rightarrow 0 $. This is  different from three-dimensional stratified turbulence where the ratio on the left hand side of (\ref{Dissipation}) is of the order of unity in the limit of strong stratification  \cite[]{Lindborg2006, Brethouwer2007}.

 The `four-fifths law' for the third order structure function of Kolmogorov turbulence is often derived using the assumption of a local cascade \cite[]{Vassilicos2015} or the assumption that dissipation stays finite in the limit of large Reynolds number \cite[]{Frisch}. As we just discussed, these assumptions seem to be violated in the case of SW wave turbulence. Nevertheless, the analogous law (\ref{eq_Kolmo}) that we derived for SW wave turbulence is satisfied with a high degree of accuracy in our simulations. In figure \ref{Flux} we wee the spectral energy flux from run W7 to the left and the third order structure function to the right. There is a broad constant flux range  with equipartition between KE and APE flux and an almost equally broad range where (\ref{eq_Kolmo}) is satisfied, with equipartition between the two third order structure functions associated with KE and APE. The assumptions of a local cascade and finite dissipation do not seem to be necessary in order for the third order structure function law  (\ref{eq_Kolmo}) to be valid.

\begin{figure}
\centerline{\includegraphics[width=5.8in]{../Pyfig/fig_flux_struct_combined}}
\caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f $. Right: Time averaged normalised third order structure functions versus $ r/L_f $. From run W7. }
\label{Flux}
\end{figure}

\begin{figure}
 \centerline{
 \includegraphics[width=3.15in]{../Pyfig/fig_spatiotempspectra}}
\caption{Frequency spectra of KE  and APE for different $ k = \mid {\bf k} \mid $, from run W7. The spectra are plotted  as functions of $\omega/\omega_l$, where
$\omega_l = c k$.  From top to bottom: $ k /\delta k
= 12$, 27, 62, 143, 327, and 746.  }
\label{fig_spatiotemp_spectra}
\end{figure}
\begin{figure}
 \centerline{
 \includegraphics[width=6.0in]{../Pyfig/fig_phys_fields_wave}}
\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
\label{Physical}
\end{figure}
\begin{figure}
\centerline{\includegraphics[width=8cm]{../Pyfig/fig_shock_sep.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance}
\end{figure}

The equipartition of KE and APE spectral fluxes suggests that the flow to leading order may be described as a collection of linear gravity waves with equipartition of KE and APE in each mode. To investigate this further we have collected time series of the flow variables from Fourier modes whose magnitude fall within certain shells,  $ \mid {\bf k} \mid \in [ k -\delta k/2, \; k+ \delta k/2] $,  and computed temporal power law spectra from these time series and averaged over all wave numbers in each shell.  In other words we have computed KE and APE frequency spectra corresponding to different magnitudes of modes. 
Figure~\ref{fig_spatiotemp_spectra} presents such
spectra plotted as a functions of the normalized frequency
$\omega/\omega_l$, where $ \omega_l = kc $ is the linear wave frequency, 
The spectra are strongly dominated by  peaks at $\omega =
\omega_l$. As can be seen, there is equipartition between KE and APE in each shell.  
Although the widening of the peaks around the linear frequency most likely is an effects of nonlinearities, 
we can quite safely conclude that the time evolution of the flow variables in each Fourier mode to leading order can be described as a linear wave.

Visualizations of the flow field give a completely different picture from what may be expected from a collection of linear gravity waves. They are totally dominated by the appearance of shock waves. In figure (\ref{Physical}) we see the divergence, $ \nabla \cdot {\bf u} $, to the left and the velocity component in the $ y $-direction, $ u_y $, to the right. The two figures at the top are from run W6 while the two figures at the bottom are from run W17. The difference between the two runs is that $ c $ is larger by a factor of $ 20 $ in W17 as compared to W6. To the left we see the shocks displayed as elongated bands of negative divergence. That the divergence is negative at the shocks can be understood from the fact that the velocity component perpendicular to the shock always has a negative jump in the direction of shock propagation. Apparently, the mean distance between the shocks is considerably smaller in run W17 with the larger value of $ c $ compared to run $ W6 $  with the smaller value of $ c $. The shocks are also visible in the figures to right where $ u_y $ is displayed. It may be interesting to note that the shocks along the $ x $-axis appear as extra sharp in this plot, since there is a strong jump  of $ u_{y} $ over these shocks, while the shocks along the $ y $-axis are hardly visible, since there is no jump of $ u_{y} $ over these.  In the two figures to the right we see a variation of the flow field over the length scale $ L_f $ which is not seen in the two figures to the left. This variation is evidently a footprint of the random forcing. 
Although the forcing length scale $ L_f $ is the same in the two simulations the mean distance, $ d $,  between the shocks is much smaller in run $ W17 $ than in run $ W6 $, which may seem a little bit surprising. We have calculated $ d $ by counting the number of negative spikes of the divergence along ten lines parallel to the $ x $-axis and ten lines parallel to the $ y $-axis and then divided  total length of all the lines by the number of spikes. Generally, we found that $ d \sim c^{-1/2} $, or
\begin{equation} \label{MeanDistance}
d \propto L_f F_f ^{1/2} \, ,
\end{equation} 
independent of resolution $ n $.
In figure (\ref{fig_distance}) we have plotted $ d/L_f $ versus Froude number for all twenty runs. As can be seen the points follow the $ F_f^{1/2} $-curve quite well, although there seems to be a slight systematic deviation at the highest $ F_f $, which may be due to the fact that $ d $ is rather close to the length of the side of the box for the largest $ F_f $.

\begin{figure}
 \centerline{
 \includegraphics[width=5.0in]{../Pyfig/fig_struct_order_246}}
\caption{From top to bottom: second, fourth and sixth order compensated and normalised longitudinal structure functions. Left: runs with $ n=3840 $: W3, W7, W11 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15. }
\label{fig_StrucFunc}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=8cm]{../Pyfig/fig_ratio_strfct}}
\caption{
Ratio of the structure functions 
$R_p(r) = \mean{|\delta u_L|^p} / \mean{|\delta u_T|^p}$ from run W4, 
for $ p = 2, \, 3, \, 4, \, 5, \, 6. $
The dotted straight lines indicate the values predicted by (\ref{Ratio}):
$R_2 = 2$, $R_3 = 3\pi/4$,  $R_4 = 8/3$, $ R_5 = 15 \pi /16 $ and $ R_6 = 16/5 $.}
\label{fig_ratio}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=12cm]{../Pyfig/fig_flatness}}
\caption{ Flatness of the longitudinal and transverse increments for
$c = 10$ and $n = 7680$. The straight continuous lines indicate the
$r^{-1}$-scaling and the straight dashed line the $r^{-3/2}$-scaling.
The inset shows the ratio $F_T/F_L$ and the corresponding value computed by the shock model, $F_T/F_L = 1.5$.  }
\label{fig_flatness}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=11cm]{../Pyfig/fig_spectra}}
\caption{Compensated and normalised energy spectra
from runs with $ n = 7680 $: W4, W8 and W15.}
\label{fig_spectra_c40}
\end{figure}

We will now present the results for the structure functions. Substituting (\ref{MeanDistance}) into (\ref{StrucFunc}) we find that the structure functions are expected to scale as
\begin{equation} \label{StrucFunc2}
\meane{|\delta u |^p}  \sim \meane{(c|\delta h |)^p} \sim  (L_f F_f^{1/2})^{p/3-1}  \eps^{p/3}  r  \, .
\end{equation}
for separation distances that fulfill the condition $ \delta x \ll r \ll d $, where $ \delta x $ is the shock width. The scaling (\ref{StrucFunc2}) is expected to become better and better fulfilled with decreasing $ c $ (or increasing $ F_f $), for two reasons. 
First, according to (\ref{MeanDistance}) $ d $ is increasing with $ F_f $, and the scaling range is therefore becoming wider for decreasing $ c $, if  $ \delta x $ or the resolution  $ n $ is kept constant. Second, according to (\ref{Strength}) the amplitude of the shocks scales as $ (\eps d)^{1/3} $, which means that the shocks are becoming increasingly strong for large $ d $ (small $ c $). Thus, the assumption that the structure functions are totally dominated by increments where a shock is crossed will be better fulfilled for small $ c $. This effect is expected to be particularly important for lower order structure functions. The scaling (\ref{StrucFunc2}) is also expected top become better and better fulfilled with increased resolution $ n $, since $ \delta x $ is decreasing with increasing $ n $.

 As expected, we have found that 
the displacement structure functions scale exactly in the same way as the velocity structure functions, why we only plot velocity structure functions. In figure (\ref{fig_StrucFunc}) we see the compensated and normalised second, fourth and sixth order longitudinal structure functions from the five simulations with $ n = 3840 $ to the left and the three simulations with $ n = 7860 $ to the right. At small separation, $ r < \delta x $, the structure functions have the dependence $ \sim r^{p} $, which is characteristic of a smooth velocity field, indicating that the shocks are reasonably well resolved. There is a small bump around the shock width, 
$ r \approx \delta x $, and at $ r > \delta x $, there is a range where the compensated structure functions are flat, or almost flat. As expected, this range is becoming broader for smaller  $ c $ (or larger $ F_f $). The range is also broader in the higher resolution runs to the right as compared to the lower resolution simulations to the left.  There is clearly a better collapse of the curves for the sixth order structure functions compared to the second order structure functions. This is not surprising, since the shock contribution to the structure functions should become increasingly dominant with increasing order. Plots of the transverse structure functions are very similar to the plots of the longitudinal structure functions, with the difference that the transverse structure functions are somewhat smoother at $ r \approx \delta x $. In figure (\ref{fig_ratio}) we see the ratio $ R_{p}(r) = \langle \mid \delta u_L \mid ^{p} \rangle /  \langle \mid \delta u_T \mid ^{p} \rangle$ for $ p= 2,3,4,5,6 $ from run W4, together with dotted straight lines, indicating the values predicted by (\ref{Ratio}). As can be seen, there is a quite good agreement with the predicted values in a limited range of separations, $ r > \delta x $. Somewhat unexpectedly, the range in which there is a good agreement becomes more narrow with increasing $ p $, an observation that we cannot explain. It may also be noted that $ R_2 \rightarrow 3 $ in the limit of small $ r $, which is the theoretical single point limit for an irrotational isotropic field. In figure (\ref{fig_flatness}) we se the flatness factor, $ F_T $,  of the transverse structure functions at the top and the ratio, $ F_T/F_L $,  at the bottom. The figures to the left show the results from the five simulations with  $ n = 3840 $ and the figures to the right show the results from the three simulations with $ n = 7860 $. The flatness factor  shows increasingly large values for decreasing $ r $ with a dependence that is rather close to but somewhat steeper than $ r^{-1} $. At small $ r $ the curves  are leveling out at very large values. For $ n = 7869 $ and $ c = 10 $ we see that $ F_{T} \approx 400 $ at small $ r $. 
The degree to which the flatness factor is increasing with decreasing scale is by some authors defined  as the most appropriate measure of spatial intermittency \cite[see for example][]{Frisch}. If this measure is used, SW wave turbulence is, indeed, extremely intermittent.  In the two figures at the bottom we see that $ F_{T}/F_{L} \approx 1.5 $ in a limited range of separations, $ r > \delta x $, which is becoming somewhat broader as $ c $ is increasing and  is also somewhat broader in the higher resolution runs shown to the right as compared to the lower resolution runs shown to the left. This is in accordance with the shock model predictions (\ref{FL}) and (\ref{FT}).

Finally, we present  energy spectra. As expected, we found that APE and KE spectra were almost indistinguishable from each other, why we only plot the total spectra. In figure (\ref{fig_spectra_c40}) we see the compensated and normalised energy spectra from the three highest resolution runs. The compensated spectra are almost flat in a broad range. In the high wave number end the compensated spectra show a `bottleneck', which we interpret as a signature of the shocks, and at the very highest wave numbers the magnitude of the spectra is strongly decreasing, indicating that the shocks are reasonably well resolved. The $ k^{-2} $-dependence and the associated scaling predicted by (\ref{Spectra}) are very well satisfied in a limited range of high wave numbers close to the bottleneck. At smaller wave number the spectra are becoming somewhat steeper than $ k^{-2} $ and the spectrum from the run with largest $ c $ does not collapse perfectly onto the other two spectra.