### 2018-10-17 Latex/ArticleNew.tex
line length

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfm}
%\documentclass[referee]{jfm} %for double spaced output for submission

\usepackage{graphicx}
\usepackage{natbib}

\usepackage{color}
\usepackage{soul}
\usepackage{booktabs}
\pdfoptionpdfminorversion 4

% \pdfcompresslevel0

% See if the author has AMS Euler fonts installed: If they have, attempt2
% to use the 'upmath' package to provide upright math.
\ifCUPmtlplainloaded \else
  \checkfont{eurm10}
  \iffontfound
    \IfFileExists{upmath.sty}
      {\typeout{^^JFound AMS Euler Roman fonts on the system,
                   using the 'upmath' package.^^J}%
       \usepackage{upmath}}
      {\typeout{^^JFound AMS Euler Roman fonts on the system, but you
                   dont seem to have the}%
       \typeout{'upmath' package installed. JFM.cls can take advantage
                 of these fonts,^^Jif you use 'upmath' package.^^J}%
       \providecommand\upi{\pi}%
      }
  \else
    \providecommand\upi{\pi}%
  \fi
\fi

% See if the author has AMS symbol fonts installed: If they have, attempt
% to use the 'amssymb' package to provide the AMS symbol characters.

\ifCUPmtlplainloaded \else
  \checkfont{msam10}
  \iffontfound
    \IfFileExists{amssymb.sty}
      {\typeout{^^JFound AMS Symbol fonts on the system, using the
                'amssymb' package.^^J}%
       \usepackage{amssymb}%
       \let\le=\leqslant  \let\leq=\leqslant
       \let\ge=\geqslant  \let\geq=\geqslant
      }{}
  \fi
\fi

% See if the author has the AMS 'amsbsy' package installed: If they have,
% use it to provide better bold math support (with \boldsymbol).

\ifCUPmtlplainloaded \else
  \IfFileExists{amsbsy.sty}
    {\typeout{^^JFound the 'amsbsy' package on the system, using it.^^J}%
     \usepackage{amsbsy}}
    {\providecommand\boldsymbol[1]{\mbox{\boldmath $##1$}}}
\fi

%%% Example macros (some are not used in this sample file) %%%

% For units of measure
\newcommand\dynpercm{\nobreak\mbox{$\;$dyn\,cm$^{-1}$}}
\newcommand\cmpermin{\nobreak\mbox{$\;$cm\,min$^{-1}$}}

% Various bold symbols
\providecommand\bnabla{\boldsymbol{\nabla}}
\providecommand\bcdot{\boldsymbol{\cdot}}
\newcommand\biS{\boldsymbol{S}}
\newcommand\etb{\boldsymbol{\eta}}

% For multiletter symbols
\newcommand\Real{\mbox{Re}} % cf plain TeX's \Re and Reynolds number
\newcommand\Imag{\mbox{Im}} % cf plain TeX's \Im
\newcommand\Rey{\mbox{\textit{Re}}}  % Reynolds number
\newcommand\Pran{\mbox{\textit{Pr}}} % Prandtl number, cf TeX's \Pr product
\newcommand\Pen{\mbox{\textit{Pe}}}  % Peclet number
\newcommand\Ai{\mbox{Ai}}            % Airy function
\newcommand\Bi{\mbox{Bi}}            % Airy function

% For sans serif characters:
% The following macros are setup in JFM.cls for sans-serif fonts in text
% and math.  If you use these macros in your article, the required fonts
% will be substitued when you article is typeset by the typesetter.
%
% \textsfi, \mathsfi   : sans-serif slanted
% \textsfb, \mathsfb   : sans-serif bold
% \textsfbi, \mathsfbi : sans-serif bold slanted (doesnt exist in CM fonts)
%
% For san-serif roman use \textsf and \mathsf as normal.
%
\newcommand\ssC{\mathsf{C}}    % for sans serif C
\newcommand\sfsP{\mathsfi{P}}  % for sans serif sloping P
\newcommand\slsQ{\mathsfbi{Q}} % for sans serif bold-sloping Q

% Hat position
\newcommand\hatp{\skew3\hat{p}}      % p with hat
\newcommand\hatR{\skew3\hat{R}}      % R with hat
\newcommand\hatRR{\skew3\hat{\hatR}} % R with 2 hats
\newcommand\doubletildesigma{\skew2\tilde{\skew2\tilde{\Sigma}}}
%       italic Sigma with double tilde

% array strut to make delimiters come out right size both ends
\newsavebox{\astrutbox}
\sbox{\astrutbox}{\rule[-5pt]{0pt}{20pt}}
\newcommand{\astrut}{\usebox{\astrutbox}}

\newcommand\GaPQ{\ensuremath{G_a(P,Q)}}
\newcommand\GsPQ{\ensuremath{G_s(P,Q)}}
\newcommand\p{\ensuremath{\partial}}
\newcommand\tti{\ensuremath{\rightarrow\infty}}
\newcommand\kgd{\ensuremath{k\gamma d}}
\newcommand\shalf{\ensuremath{{\scriptstyle\frac{1}{2}}}}
\newcommand\sh{\ensuremath{^{\shalf}}}
\newcommand\smh{\ensuremath{^{-\shalf}}}
\newcommand\squart{\ensuremath{{\textstyle\frac{1}{4}}}}
\newcommand\thalf{\ensuremath{{\textstyle\frac{1}{2}}}}
\newcommand\Gat{\ensuremath{\widetilde{G_a}}}
\newcommand\ttz{\ensuremath{\rightarrow 0}}
\newcommand\ndq{\ensuremath{\frac{\mbox{$\partial$}}{\mbox{$\partial$} n_q}}}
\newcommand\sumjm{\ensuremath{\sum_{j=1}^{M}}}
\newcommand\pvi{\ensuremath{\int_0^{\infty}%
  \mskip \ifCUPmtlplainloaded -30mu\else -33mu\fi -\quad}}

\newcommand\etal{\mbox{\textit{et al.}}}
\newcommand\etc{etc.\ }
\newcommand\eg{e.g.\ }

\newcommand{\uu}{\textbf{u}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\JJ}{\textbf{J}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ff}{\textbf{f}}
\newcommand{\FF}{\textbf{F}}
\newcommand{\baa}{\textbf{a}}
\newcommand{\bb}{\textbf{b}}
\newcommand{\NN}{\textbf{N}}

\newcommand{\eps}{\varepsilon}

\newcommand{\ErtelPV}{\mathcal{Q}}

\newcommand{\fdiss}{f_{\mbox{\scriptsize diss}}}

\newcommand{\D}{\mbox{D}}

\newcommand{\eez}{\boldsymbol{e_z}}
\newcommand{\scalarprod}[2]{\big( #1 \, , \ #2 \big)_{\kk}}

\newcommand{\mean}[1]{\langle #1 \rangle}

\newcommand{\meane}[1]{\langle #1 \rangle}
% \newcommand{\meane}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}e}}}

\newcommand{\meanx}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\mbox{\scriptsize$\xx$}}}}

\newcommand{\meant}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\theta}}}

\newcommand{\means}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}s}}}

\newcommand{\shocks}{ {\mbox{\tiny shocks}} }

\newcommand{\kmax}{k_{\mbox{\scriptsize max}}}
\newcommand{\kdiss}{k_{\mbox{\scriptsize diss}}}

\newcommand{\mA}{\mathcal{A}}

\newlength{\halfwidth}
\setlength{\halfwidth}{2.6in}

\newcommand{\PA}[1]{{\color{green}#1}}

\newcommand{\Add}[1]{{\color{blue}#1}}
% \newcommand{\Add}[1]{#1}
\newcommand{\Remove}[1]{{\color{red}\st{#1}}}
% \newcommand{\Remove}[1]{}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title[Shallow water wave turbulence]%
{ Shallow water wave turbulence}

\author[P. Augier, A.V. Mohanan and E. Lindborg]%
{Pierre Augier$^{1}$,   A. V. Mohanan$^{2}$
and Erik Lindborg$^2$ \ns }

\affiliation{$^2$ KTH Mechanics,
SE-100 44 Stockholm, Sweden\\[\affilskip]
$^1$ LEGI, BP53,
38041 Grenoble Cedex, France}

\pubyear{2013}
\volume{???}
\pagerange{???--???}
% Do not enter received and revised dates. These will be entered by
% the editorial office.
\date{?; revised ?; accepted ?. - To be entered by editorial office}
%\setcounter{page}{1}
\begin{document}

\maketitle

% \baselineskip 8mm

\begin{abstract}

The dynamics of irrotational shallow water wave turbulence forced in large
scales and dissipated at small scales is investigated. First, we derive the
shallow water analogue of the `four-fifths law' of three-dimensional turbulence
for a third order structure function involving velocity and displacement
increments. Using this relation and assuming that the flow is dominated by
shocks we develop a simple model predicting that the shock amplitude scales as
$ (\epsilon d)^{1/3} $, where $ \eps $ is the mean dissipation rate and $ d $
the mean distance between the shocks, and that the $ p $:th order displacement
and velocity structure functions scale as $ (\eps d)^{p/3} r/d $, where $ r $
is the separation. Then we carry out a series of forced simulations with
resolutions up to $7680^2$, varying the Froude number, $F_{f} = \eps^{1/3} /
ck_f^{1/3} $, where $ k_f $ is the forcing wave number and $c$ is the wave
speed. In all simulations a stationary state is reached in which there is a
constant spectral energy flux and equipartition between kinetic and potential
energy in the constant flux range. The third order structure function relation
is satisfied with a high degree of accuracy. Mean energy is found to scale as $
E \sim \sqrt{\epsilon c/k_f} $, and is also dependent on resolution, indicating
that shallow water wave turbulence does not fit into the paradigm of a
Richardson-Kolmogorov cascade. In all simulations shocks develop, displayed as
long thin bands of negative divergence in flow visualizations. The mean
distance between the shocks is found to scale as $ d \sim F_f^{1/2}/k_f $.
Structure functions of second and higher order are found to scale in good
agreement with the model. We conclude that in the weak limit, $ F_f \rightarrow
0 $, shocks will become denser and weaker and finally disappear for a finite
Reynolds number. On the other hand, for a given $ F_{f} $, no matter how small,
shocks will prevail if the Reynolds number is sufficiently large.

\end{abstract}

% \begin{keywords}
% \end{keywords}

\section{Introduction}

The shallow water (SW) equations  have been widely used to study basic
mechanisms occurring in geophysical flows \cite[see for
example][]{VallisLIVRE2006}.
%
The equations are based on the hydrostatic approximation which is
extremely well satisfied for a very wide range of scales in the oceans
and in the atmosphere.  They also rely on a stronger hypothesis
regarding the density stratification.  The fluid is assumed to be
structured in a limited number of thin homogeneous layers, which leads
to equations involving only two-dimensional operators. The two-layer
SW equations capture the baroclinic instability which is of
primary importance for the dynamics of the oceans and of the
atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
%
The single layer SW equations are very similar to
the two-dimensional compressible Navier-Stokes equations.  The
equations do not capture baroclinic instability but they constitute
one of the simplest hydrodynamic models with coexisting eigenmodes of
the linearized operator with zero linear frequency and with finite
non-zero linear frequency, $\omega  \pm\sqrt{f^2 + c^2
k^2}$, where $f$ is the Coriolis parameter, $c$ the gravity wave speed and $k
= |\kk|$ the wave number.
%

For geophysical applications, it is convenient to consider the limit
for which the linear wave frequency  is much larger than the
characteristic nonlinear frequency $ U k$, where $U$ is a
characteristic velocity.  In this limit, one can obtain a quasi-geostrophic
single layer SW model, i.e. derive self-consistent evolution
equations for a zero frequency variable, the Charney potential
vorticity, which is totally decoupled from the fast waves.
%
For finite rotation and wave speed, there are slow balanced motions,
which are associated with relatively weak non-zero linear frequency
modes and there is only a weak
interaction between waves and the balanced motions.
%
Due to their relative simplicity, the SW equations have
been used to study issues like the projection on the balanced manifold
\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
waves by balanced flows \cite[]{FargeSadourny1989,
LahayeZeitlin2012,Vanneste2013}.
In this study, we consider the
same equations in order to address a related but different issue.

Balanced flows have a dynamics which is very similar to
two-dimensional turbulence, with an upscale energy cascade. %
Small scale energy dissipation vanishes in the limit of zero
viscosity.
%
In contrast to this, statistical mechanics indicates that wave energy
should be transferred towards small scales \cite[]{Warn1986}.
Numerical simulations have shown that long waves lose their energy by
transferring it to shorter waves \cite[]{Sadourny1975,
FargeSadourny1989, YuanHamilton1994}.  The mechanism of a wave cascade
has been proposed to explain the attraction towards the balanced flows
\cite[]{Sadourny1975}.  \cite{YuanHamilton1994} showed that
statistically stationary shallow-water flows can be obtained forcing
quasi-geostrophic modes at large scales and with dissipation only at
small scales.
%
They propose that the $k^{-5/3}$ spectra that are observed at the
atmospheric meso-scales  \cite[]{NastromGage1985,LiLindborg2018}
and reproduced by some atmospheric models based on the primitive
equations \cite[]{KoshykHamilton2001, Skamarock2004,
Hamilton_etal2008} could be reproduced by the one-layer
shallow-water model by a downscale energy cascade of waves.

However, few studies have investigated the dynamics of the downscale
energy cascade of waves in SW models.
%
Most numerical studies have investigated decaying flows and have not
focused on the cascade \cite[]{FargeSadourny1989,
LarichevMcWilliams1991, SpallMcWilliams1992,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
%
Moreover, there are strong discrepancies between different numerical
results and between these results and available theoretical results.
%
\cite{FargeSadourny1989} reported spectra scaling as $k^{-4}$
whereas those shown by \cite{LahayeZeitlin2012} scale as $k^{-6}$.

Some predictions are based on the weak wave turbulence formalism
\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011}. In the case of
inertia-gravity waves, \cite{FalkovichMedvedev1992} showed that the exact
solutions of the kinetic equation corresponding to a constant downscale energy
flux is associated with a spectrum scaling as $E(k) \sim k^{-8/3}$, while
\cite{ZakharovSagdeev1970} used weak turbulence formalism to derive a spectrum
of the form $ E(k) \sim k^{-3/2} $ for acoustic turbulence. This prediction is
of particular interest, since the aqoustic equations and the SW equations are
identical when a first order expansion is made in density or displacement
pertubations. It may be questioned, however, that weak turbulence formalism is
applicable to acoustic and SW wave turbulence, no matter how weak the
nonlinearities are supposed to be. As pointed out by \cite{KadomtsevPet1970}
and \cite{FalkovichMeyer1996}, when dispersion is small enough, waves moving in
the same direction have the same velocity and interact strongly, which leads to
the formation of shocks. In this case, the weak wave turbulence formalism is
unable to give valuable results. Such sharp and long structures have been
observed in many simulations \cite[]{FargeSadourny1989,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012} but their effects are not
well understood.
%
As shown by \cite{Burgers1948} the energy spectrum of a flow dominated by
shocks will scale as $ k^{-2} $, a prediction which has been confirmed by a
great number of simulations of Burgers equation \cite[for a review, see ]
[]{Frisch2001}. The shocks also lead to very strong intermittency, as
demonstrated by \cite{BouchaudMezardParisi1995} and
\cite{WeinanKhaninMazelSinai1997} for Burgers turbulence.
%
\cite{FalkovichMeyer1996} simulated acoustic turbulence dominated by shocks
and reported spectra scaling as $ k^{-2} $.
\cite{YuanHamilton1994} carried out
forced dissipative simulations of the SW equations showing similar spectra.
However, the resolution in both these studies was quite coarse, ($386^2$
and $101^2$ grid points, respectively) so that the scaling range was very
narrow.  Moreover, \cite{FalkovichMeyer1996} used a very particular
type of strongly anisotropic forcing with waves in only one
direction while \cite{YuanHamilton1994} forced in geostrophic modes
which severely complicates the interpretation.

The dynamics of waves in complex SW flows is not well
understood.  Even wave dynamics in absence of a balanced flow is not well
understood. A statistically stationary wave cascade has never been
simulated at high resolutions.
%
Is it possible to obtain a statistically stationary state in the very
fast wave regime, and how does the mean energy in this hypothetical
regime depends on the forcing and the wave speed?
% What is the slope of
How do energy spectra scale with the input parameters? From a theoretical point
of view, the most interesting issue is, perhaps, the nature of the dynamics in
the limit of weak nonlinearities. Will shocks prevail in this limit?
%
Such questions are still open and straightforward simulations may
provide the answers.
In this paper, we investigate the dynamics of single layer SW wave turbulence
forced  at large scales by means of
simulations with up the  $7680^2$ grid points.
%
Before presenting the results of the simulations we develop some theory.
%

\section{Theory}
\label{section_basic_theory}
\input{section_basic_theory}

\label{section_advanced_theory}
\input{section_advanced_theory}

\section{Simulations}
\label{section_methods}
\input{section_methods}

%\label{section_wave_cascade_f=0}
%\input{section_wave_cascade_f=0}

%\section{Effects of global rotation}
%\label{section_rotation}
%\input{section_rotation}

\section{Conclusions}

We have carried out a theoretical and numerical investigation of SW wave
turbulence. First, we derived the SW analogue (\ref{eq_Kolmo}) of the the
`four-fifths' law of Kolmogorov turbulence. Using this relation and
straightforward statistical and geometrical arguments we developed a simple
shock model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3}
$, where $ d $ is the mean distance between the shocks, and that the $ p $:th
order structure function above a certain minimum will scale as $ (\epsilon
d)^{p/3} r/d $. Then, we carried out a series of twenty forced dissipative
simulations, varying the Froude number and the resolution. In all simulations,
a statistically stationary state were reached and in all simulations we
observed shocks. The flow variables in each Fourier mode were found to evolve
in accordance with linear wave dynamics, with equipartition between APE and KE
over a period. The third order structure function relation was fulfilled with a
high degree of accuracy. From the simulations we made two interesting
observations that fall outside the predictive range of the model. The first
observation was that mean energy in the stationary state approximately scales
as (\ref{MeanEnergy}), suggesting that the normalised dissipation
(\ref{Dissipation}) will go to zero both in the limit of small viscosity and
large wave speed. This is different from three dimensional Kolmogorov
turbulence as well as strongly stratified turbulence, where, in both cases,
dissipation is finite in the limit of small viscosity, and in the latter case
also is independent of the degree of stratification. This observation suggests
that SW wave turbulence does not fit into the paradigm of a local
Richardson-Kolmogorov cascade. The second observation we made was that the mean
distance between the shocks scales as $ c^{-1/2} $, if the other parameters are
held constant. As we increased $ c $, the shocks were thus becoming denser.
Using this observation we tested the shock model by plotting the normalised and
compensated structure functions of different orders. Generally we found that
there was a quite good agreement between the simulation results and the model
predictions, becoming better with decreasing wave speed and increasing
resolution.

Combining the model and the observations we can get a quite good picture of the
dynamics in the limit of weak nonlinearities, $ F_{f} \rightarrow 0 $. The
picture may become more vivid if we say `wave breaking' were we previously have
talked about shock formation, and also say `wave lengths' were we previously
have talked about length scales. We identify the longest wave length, $
\lambda_l $, with the forcing scale $ L_f $ and the breaking wave length, $
\lambda_b $, with the mean distance, $ d $, between the shocks. There are two
dynamically different regimes, the short wave regime, $ [\delta x, \;
\lambda_b] $, and the long wave regime, $ [\lambda_b, \, \lambda_l] $. We first
consider the latter, whose width scales $ \lambda_{l} /\lambda_b \sim
F_f^{-1/2} $. Identifying the amplitude of the breaking waves with the shock
amplitude, $ (\eps d)^{1/3} $, we find that the ratio between the amplitude of
the longest waves and the breaking waves scales as $ E^{1/2}/(\eps d)^{1/3}
\sim F_{f}^{-5/12} n^{\alpha/2} $, where we may substitute $ n $ with a
Reynolds number to some power. Thus, for small $ F_f $ and large Reynolds
number the longest waves are not directly influenced by wave breaking. The
Reynolds number dependence of the amplitude ratio makes it unlikely that a
general similarity theory can be easily formulated for this range. Therefore,
predictions from weak turbulence theory \cite[for
example][]{ZakharovSagdeev1970} are not likely to hold for this range, although
it is not directly influenced by wave breaking. The exact values of the power
law exponents in the argument leading to this conclusion are, of course, not
important. As for the short wave regime, the width scales as $ \lambda_b
/\delta x \sim F_f^{1/2} $ for fixed $ \delta x $ or fixed Reynolds number. As
$ F_{f} \rightarrow 0 $ the range will eventually become so narrow so that
breaking will disappear. On the other hand, it may be argued that wave breaking
will take place for any small $ F_f $, provided that the Reynolds number is
sufficiently large, since the shock width $ \delta x $ is decreasing
indefinitely with increasing Reynolds number.

The question that originally motivated this study was if SW wave turbulence can
be used as a model for energy dissipation in the the atmosphere and the oceans.
Undoubtedly, the answer to this question is negative and the reason for this is
not primarily the appearance of shocks and the associated $ k^{-2} $-spectrum
which is not consistent with the observed $ k^{-5/3} $-spectrum. The main
reason is that SW wave turbulence is a far too inefficient agent of
dissipation. Since the Reynolds number of geophysical flows is extremely large
and the stratification is strong, the relation (\ref{Dissipation}) implies that
dissipation would be absent in the interior of the atmosphere and the oceans
with SW turbulence as the responsible agent. This not what we observe.

Although the answer to the question that originally motivated this study is
negative, it may lead us or some other investigators to pursue another
interesting research direction. As pointed out in the introduction, the
acoustic equations are to the lowest order expansion in density pertubations
identical to the SW equations. This study may therefore give important clues to
the long standing problem of acoustic turbulence, in two and three dimensions.
A similar shock model as we developed in two dimensions may easily be
formulated in three dimensions, where the third order structure function law
will be similar to (\ref{eq_Kolmo}), with the only difference that the constant
$ 4 $ on the right hand side should be replaced by $ 8/3 $. In three dimensions
we find
\begin{equation}
\meane{ |\delta \uu|^2 \delta J_L }
+ c^2\meane{ (\delta h)^2 \delta u_L } = - \frac{8}{3} \varepsilon r. \label{eq_Kolmo3}
\end{equation}
In the case of acoustic turbulence $ \delta h $ should be interpreted as a
normalised density increment rather than a displacement increment. If shocks in
three dimensions are assumed to form as smooth surfaces rather than smooth
lines, then the mean distance, $ d $, between the shocks will appear in the
model as $ d \sim {\cal{V}}/A_s $, where $ A_{s} $ is the total shock area and
$ \cal{V} $ is the volume of the domain. The shock amplitude will thus scale as
$ (\epsilon d)^{1/3} $ and the structure functions will scale as $ (\epsilon
d)^{p/3} r/d $ also in three dimensions, provided, of course, that shocks will
form in a similar way as in two dimensions. If they do, the most interesting
questions are if the mean energy and the mean distance between the shocks scale
in a similar way in three dimensions as in two dimensions. Simulations or
experiments of acoustic turbulence may provide answers to these questions --
answers that will lead to further clues to a theoretical understanding of
acoustic turbulence. We hope that the present study will stimulate research in
this direction.

%\appendix

%\section{Kinetic energy dissipation and forcing rates}
%\label{app_comp}
%\input{appendix_dissip_forcing}

\bibliographystyle{jfm}
\bibliography{biblio}

\end{document}