### 2018-10-23 Latex/article.tex
Update Erik's files

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfm}
%\documentclass[referee]{jfm} %for double spaced output for submission

\usepackage{graphicx}
\usepackage{natbib}

\usepackage{color}
\usepackage{soul}
\usepackage{booktabs}
\pdfoptionpdfminorversion 4

% \pdfcompresslevel0

% See if the author has AMS Euler fonts installed: If they have, attempt2
% to use the 'upmath' package to provide upright math.
\ifCUPmtlplainloaded \else
  \checkfont{eurm10}
  \iffontfound
    \IfFileExists{upmath.sty}
      {\typeout{^^JFound AMS Euler Roman fonts on the system,
                   using the 'upmath' package.^^J}%
       \usepackage{upmath}}
      {\typeout{^^JFound AMS Euler Roman fonts on the system, but you
                   dont seem to have the}%
       \typeout{'upmath' package installed. JFM.cls can take advantage
                 of these fonts,^^Jif you use 'upmath' package.^^J}%
       \providecommand\upi{\pi}%
      }
  \else
    \providecommand\upi{\pi}%
  \fi
\fi

% See if the author has AMS symbol fonts installed: If they have, attempt
% to use the 'amssymb' package to provide the AMS symbol characters.

\ifCUPmtlplainloaded \else
  \checkfont{msam10}
  \iffontfound
    \IfFileExists{amssymb.sty}
      {\typeout{^^JFound AMS Symbol fonts on the system, using the
                'amssymb' package.^^J}%
       \usepackage{amssymb}%
       \let\le=\leqslant  \let\leq=\leqslant
       \let\ge=\geqslant  \let\geq=\geqslant
      }{}
  \fi
\fi

% See if the author has the AMS 'amsbsy' package installed: If they have,
% use it to provide better bold math support (with \boldsymbol).

\ifCUPmtlplainloaded \else
  \IfFileExists{amsbsy.sty}
    {\typeout{^^JFound the 'amsbsy' package on the system, using it.^^J}%
     \usepackage{amsbsy}}
    {\providecommand\boldsymbol[1]{\mbox{\boldmath $##1$}}}
\fi

%%% Example macros (some are not used in this sample file) %%%

% For units of measure
\newcommand\dynpercm{\nobreak\mbox{$\;$dyn\,cm$^{-1}$}}
\newcommand\cmpermin{\nobreak\mbox{$\;$cm\,min$^{-1}$}}

% Various bold symbols
\providecommand\bnabla{\boldsymbol{\nabla}}
\providecommand\bcdot{\boldsymbol{\cdot}}
\newcommand\biS{\boldsymbol{S}}
\newcommand\etb{\boldsymbol{\eta}}

% For multiletter symbols
\newcommand\Real{\mbox{Re}} % cf plain TeX's \Re and Reynolds number
\newcommand\Imag{\mbox{Im}} % cf plain TeX's \Im
\newcommand\Rey{\mbox{\textit{Re}}}  % Reynolds number
\newcommand\Pran{\mbox{\textit{Pr}}} % Prandtl number, cf TeX's \Pr product
\newcommand\Pen{\mbox{\textit{Pe}}}  % Peclet number
\newcommand\Ai{\mbox{Ai}}            % Airy function
\newcommand\Bi{\mbox{Bi}}            % Airy function

% For sans serif characters:
% The following macros are setup in JFM.cls for sans-serif fonts in text
% and math.  If you use these macros in your article, the required fonts
% will be substitued when you article is typeset by the typesetter.
%
% \textsfi, \mathsfi   : sans-serif slanted
% \textsfb, \mathsfb   : sans-serif bold
% \textsfbi, \mathsfbi : sans-serif bold slanted (doesnt exist in CM fonts)
%
% For san-serif roman use \textsf and \mathsf as normal.
%
\newcommand\ssC{\mathsf{C}}    % for sans serif C
\newcommand\sfsP{\mathsfi{P}}  % for sans serif sloping P
\newcommand\slsQ{\mathsfbi{Q}} % for sans serif bold-sloping Q

% Hat position
\newcommand\hatp{\skew3\hat{p}}      % p with hat
\newcommand\hatR{\skew3\hat{R}}      % R with hat
\newcommand\hatRR{\skew3\hat{\hatR}} % R with 2 hats
\newcommand\doubletildesigma{\skew2\tilde{\skew2\tilde{\Sigma}}}
%       italic Sigma with double tilde

% array strut to make delimiters come out right size both ends
\newsavebox{\astrutbox}
\sbox{\astrutbox}{\rule[-5pt]{0pt}{20pt}}
\newcommand{\astrut}{\usebox{\astrutbox}}

\newcommand\GaPQ{\ensuremath{G_a(P,Q)}}
\newcommand\GsPQ{\ensuremath{G_s(P,Q)}}
\newcommand\p{\ensuremath{\partial}}
\newcommand\tti{\ensuremath{\rightarrow\infty}}
\newcommand\kgd{\ensuremath{k\gamma d}}
\newcommand\shalf{\ensuremath{{\scriptstyle\frac{1}{2}}}}
\newcommand\sh{\ensuremath{^{\shalf}}}
\newcommand\smh{\ensuremath{^{-\shalf}}}
\newcommand\squart{\ensuremath{{\textstyle\frac{1}{4}}}}
\newcommand\thalf{\ensuremath{{\textstyle\frac{1}{2}}}}
\newcommand\Gat{\ensuremath{\widetilde{G_a}}}
\newcommand\ttz{\ensuremath{\rightarrow 0}}
\newcommand\ndq{\ensuremath{\frac{\mbox{$\partial$}}{\mbox{$\partial$} n_q}}}
\newcommand\sumjm{\ensuremath{\sum_{j=1}^{M}}}
\newcommand\pvi{\ensuremath{\int_0^{\infty}%
  \mskip \ifCUPmtlplainloaded -30mu\else -33mu\fi -\quad}}

\newcommand\etal{\mbox{\textit{et al.}}}
\newcommand\etc{etc.\ }
\newcommand\eg{e.g.\ }

\newcommand{\uu}{\textbf{u}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\JJ}{\textbf{J}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ff}{\textbf{f}}
\newcommand{\FF}{\textbf{F}}
\newcommand{\baa}{\textbf{a}}
\newcommand{\bb}{\textbf{b}}
\newcommand{\NN}{\textbf{N}}

\newcommand{\ErtelPV}{\mathcal{Q}}

\newcommand{\fdiss}{f_{\mbox{\scriptsize diss}}}

\newcommand{\D}{\mbox{D}}

\newcommand{\eez}{\boldsymbol{e_z}}
\newcommand{\scalarprod}[2]{\big( #1 \, , \ #2 \big)_{\kk}}

\newcommand{\mean}[1]{\langle #1 \rangle}

\newcommand{\meane}[1]{\langle #1 \rangle}
% \newcommand{\meane}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}e}}}

\newcommand{\meanx}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\mbox{\scriptsize$\xx$}}}}

\newcommand{\meant}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\theta}}}

\newcommand{\means}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}s}}}

\newcommand{\shocks}{ {\mbox{\tiny shocks}} }

\newcommand{\kmax}{k_{\mbox{\scriptsize max}}}
\newcommand{\kdiss}{k_{\mbox{\scriptsize diss}}}

\newcommand{\mA}{\mathcal{A}}

\newlength{\halfwidth}
\setlength{\halfwidth}{2.6in}

\newcommand{\PA}[1]{{\color{green}#1}}

\newcommand{\Add}[1]{{\color{blue}#1}}
% \newcommand{\Add}[1]{#1}
\newcommand{\Remove}[1]{{\color{red}\st{#1}}}
% \newcommand{\Remove}[1]{}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title[Shallow water wave turbulence]%
{ Shallow water wave turbulence}

\author[P. Augier, A.V. Mohanan and E. Lindborg]%
{Pierre Augier$^{1}$,   A. V. Mohanan$^{2}$
and Erik Lindborg$^2$ \ns }

\affiliation{$^2$ KTH Mechanics,
SE-100 44 Stockholm, Sweden\\[\affilskip]
$^1$ LEGI, BP53,
38041 Grenoble Cedex, France}

\pubyear{2013}
\volume{???}
\pagerange{???--???}
% Do not enter received and revised dates. These will be entered by
% the editorial office.
\date{?; revised ?; accepted ?. - To be entered by editorial office}
%\setcounter{page}{1}
\begin{document}

\maketitle

% \baselineskip 8mm

\begin{abstract}

The dynamics of irrotational shallow water wave turbulence forced in large
scales and dissipated at small scales is investigated. First, we derive the
shallow water analogue of the `four-fifths law' of Kolmogorov turbulence
for a third order structure function involving velocity and displacement
increments. Using this relation and assuming that the flow is dominated by
shocks we develop a simple model predicting that the shock amplitude scales as
$ (\epsilon d)^{1/3} $, where $ \epsilon $ is the mean dissipation rate and $ d $
the mean distance between the shocks, and that the $ p $:th order displacement
and velocity structure functions scale as $ (\epsilon d)^{p/3} r/d $, where $ r $
is the separation. Then we carry out a series of forced simulations with
resolutions up to $7680^2$, varying the Froude number, $F_{f} = \epsilon^{1/3} /
ck_f^{1/3} $, where $ k_f $ is the forcing wave number and $c$ is the wave
speed. In all simulations a stationary state is reached in which there is a
constant spectral energy flux and equipartition between kinetic and potential
energy in the constant flux range. The third order structure function relation
is satisfied with a high degree of accuracy. Mean energy is found to scale as $
E \sim \sqrt{\epsilon c/k_f} $, and is also dependent on resolution, indicating
that shallow water wave turbulence does not fit into the paradigm of a
Richardson-Kolmogorov cascade. In all simulations shocks develop, displayed as
long thin bands of negative divergence in flow visualisations. The mean
distance between the shocks is found to scale as $ d \sim F_f^{1/2}/k_f $.
Structure functions of second and higher order are found to scale in good
agreement with the model. We conclude that in the weak limit, $ F_f \rightarrow
0 $, shocks will become denser and weaker and finally disappear for a finite
Reynolds number. On the other hand, for a given $ F_{f} $, no matter how small,
shocks will prevail if the Reynolds number is sufficiently large.

\end{abstract}

% \begin{keywords}
% \end{keywords}

\section{Introduction}

The shallow water (SW) equations  have been widely used to study basic
mechanisms occurring in geophysical flows \cite[see for
example][]{VallisLIVRE2006}.
%
The equations are based on the hydrostatic approximation which is
extremely well satisfied for a very wide range of scales in the oceans
and in the atmosphere.  They also rely on a stronger hypothesis
regarding the density stratification.  The fluid is assumed to be
structured in a limited number of thin homogeneous layers, which leads
to equations involving only two-dimensional operators. The two-layer
SW equations capture the baroclinic instability which is of
primary importance for the dynamics of the oceans and of the
atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
%
The single layer SW equations are very similar to
the two-dimensional compressible Navier-Stokes equations.  The
equations do not capture baroclinic instability but they constitute
one of the simplest hydrodynamic models with coexisting eigenmodes of
the linearized operator with zero linear frequency and with finite
non-zero linear frequency, $\omega = \pm\sqrt{f^2 + c^2
k^2}$, where $f$ is the Coriolis parameter, $c$ the gravity wave speed and $k
= |\kk|$ the wave number.
%

For geophysical applications, it is convenient to consider the limit
for which the linear wave frequency  is much larger than the
characteristic nonlinear frequency $ U k$, where $U$ is a
characteristic velocity.  In this limit, one can obtain a quasi-geostrophic
single layer SW model, i.e. derive self-consistent evolution
equations for a zero frequency variable, the Charney potential
vorticity, which is totally decoupled from the fast waves.
%
For finite rotation and wave speed, there are slow balanced motions,
which are associated with relatively weak non-zero linear frequency
modes and there is only a weak
interaction between waves and the balanced motions.
%
Due to their relative simplicity, the SW equations have
been used to study issues like the projection on the balanced manifold
\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
waves by balanced flows \cite[]{FargeSadourny1989,
LahayeZeitlin2012,Vanneste2013}.
In this study, we consider the
same equations in order to address a related but different issue.

Balanced flows have a dynamics which is very similar to
two-dimensional turbulence, with an upscale energy cascade. %
Small scale energy dissipation vanishes in the limit of zero
viscosity.
%
In contrast to this, statistical mechanics indicates that wave energy
should be transferred towards small scales \cite[]{Warn1986}.
Numerical simulations have shown that long waves lose their energy by
transferring it to shorter waves \cite[]{Sadourny1975,
FargeSadourny1989, YuanHamilton1994}.  The mechanism of a wave cascade
has been proposed to explain the attraction towards the balanced flows
\cite[]{Sadourny1975}.  \cite{YuanHamilton1994} showed that
statistically stationary shallow-water flows can be obtained forcing
quasi-geostrophic modes at large scales and with dissipation only at
small scales.
%
They propose that the $k^{-5/3}$ spectra that are observed at the
atmospheric meso-scales  \cite[]{NastromGage1985,LiLindborg2018}
and reproduced by some atmospheric models based on the primitive
equations \cite[]{KoshykHamilton2001, Skamarock2004,
Hamilton_etal2008} could be reproduced by the one-layer
shallow-water model by a downscale energy cascade of waves.
However, few studies have investigated the dynamics of the downscale
energy cascade of waves in SW models.
Most numerical studies have investigated decaying flows and have not
focused on the cascade \cite[]{FargeSadourny1989,
LarichevMcWilliams1991, SpallMcWilliams1992,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
%
Moreover, there are strong discrepancies between different numerical
results and between these results and available theoretical results.
%
\cite{FargeSadourny1989} reported spectra scaling as $k^{-4}$
whereas those shown by \cite{LahayeZeitlin2012} scale as $k^{-6}$.

Some predictions are based on the weak wave turbulence formalism
\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011}. In the case of
inertia-gravity waves, \cite{FalkovichMedvedev1992} showed that the exact
solutions of the kinetic equation corresponding to a constant downscale energy
flux is associated with a spectrum scaling as $E(k) \sim k^{-8/3}$, while
\cite{ZakharovSagdeev1970} used weak turbulence formalism to derive a spectrum
of the form $ E(k) \sim k^{-3/2} $ for acoustic turbulence. This prediction is
of particular interest, since the acoustic equations and the SW equations are
identical when a first order expansion is made in density or displacement
perturbations. It may be questioned, however, that weak turbulence formalism is
applicable to acoustic and SW wave turbulence, no matter how weak the
nonlinearities are supposed to be. As pointed out by \cite{KadomtsevPet1970}
and \cite{FalkovichMeyer1996}, when dispersion is small enough, waves moving in
the same direction have the same velocity and interact strongly, which leads to
the formation of shocks. In this case, the weak wave turbulence formalism is
unable to give valuable results. Such sharp and long structures have been
observed in many simulations \cite[]{FargeSadourny1989,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012} but their effects are not
well understood.
%
As shown by \cite{Burgers1948} the energy spectrum of a flow dominated by
shocks will scale as $ k^{-2} $, a prediction which has been confirmed by a
great number of simulations of Burgers equation \cite[for a review, see ]
[]{Frisch2001}. The shocks also lead to very strong intermittency, as
demonstrated by \cite{BouchaudMezardParisi1995} and
\cite{WeinanKhaninMazelSinai1997} for Burgers turbulence.
%
\cite{FalkovichMeyer1996} simulated acoustic turbulence dominated by shocks
and reported spectra scaling as $ k^{-2} $.
\cite{YuanHamilton1994} carried out
forced dissipative simulations of the SW equations showing similar spectra.
However, the resolution in both these studies was quite coarse, ($386^2$
and $101^2$ grid points, respectively) so that the scaling range was very
narrow.  Moreover, \cite{FalkovichMeyer1996} used a very particular
type of strongly anisotropic forcing with waves in only one
direction while \cite{YuanHamilton1994} forced in geostrophic modes
which severely complicates the interpretation.

The dynamics of waves in complex SW flows is not well
understood.  Even wave dynamics in absence of a balanced flow is not well
understood. A statistically stationary wave cascade has never been
simulated at high resolutions.
%
Is it possible to obtain a statistically stationary state in the very
fast wave regime, and how does the mean energy in this hypothetical
regime depends on the forcing and the wave speed?
% What is the slope of
How do energy spectra scale with the input parameters? From a theoretical point
of view, the most interesting issue is, perhaps, the nature of the dynamics in
the limit of weak nonlinearities. Will shocks prevail in this limit?
%
Such questions are still open and straightforward simulations may
provide the answers.
In this paper, we investigate the dynamics of single layer SW wave turbulence
forced  at large scales by means of
simulations with up the  $7680^2$ grid points.
%
Before presenting the results of the simulations we develop some theory.
%

\section{Theory}

In a non-rotating system of reference the SW equations for a thin layer of fluid over a flat bottom and with a free
surface can be written as  \cite[see for example][]{VallisLIVRE2006}
\begin{eqnarray} \label{UEq}
\p_t \uu + \uu\cdot \bnabla  \uu   
&=& - c^2 \bnabla h + \frac{\nu}{h}  \bnabla^2 \uu,   \label{eq_uu}\\ \label{HEq}
\p_t   h    &=& - \bnabla \cdot (h \uu).  \label{eq_h}
\end{eqnarray}
where $\uu$ is the horizontal velocity, %
$h = 1 + \eta $ the non-dimensional thickness of the fluid layer,  $ \eta $ being the surface displacement, 
$\nu$ the kinematic viscosity and $ c $ is the gravity wave speed. The SW equations conserve mass, $ h $, and momentum $ {\bf J} = h{\bf u} $. As a matter of fact, we have constructed the viscous term in (\ref{UEq}) in such a way that momentum is conserved and energy dissipation is positive definite. This can only be accomplished by permitting the term to be nonlinear in the flow variables. There are different forms of nonlinear terms that fulfil both conditions and it may be disputed which is the best one. However, in the simulations we will use diffusion terms in both (\ref{UEq}) and (\ref{HEq}) that will not fulfil any of these two conditions in a strict sense and we will therefore not discuss this matter further. The inviscid SW equations conserve a local
quantity along trajectories of the fluid particles, the Ertel
potential vorticity $\ErtelPV = \zeta /h$, where $\zeta  $ is the vorticity, which in a rotating system should be interpreted as absolute vorticity.  This conservation law will be of no relevance in the present study in which we only consider
irrotational flows for which the velocity can be written as $ \uu = \bnabla \phi $ , where $ \phi $ is the velocity potential. In a domain with no net fluxes through the boundaries the inviscid equations conserve energy, $ E = E_K + E_P $, where $ E_K =   {\bf J}  \cdot {\uu} /2 $ is kinetic energy and $ E_P = c^{2} h^2/2 $ is potential energy. The latter can be split into three terms as \begin{equation}
E_P = c^2/2 + c^2\eta+ c^2 \eta^2/2, 
\end{equation}
where the first term corresponds to the potential energy of a state with no
surface displacement.  The third term, which is quadratic in the surface displacement, may be named available potential energy (APE),
$E_A = c^2\eta^2/2$, a concept introduced by  \cite{Lorenz1955}. As a consequence of mass and energy conservation the SW equations also conserve the sum of kinetic and available potential energy, $ E_K + E_A $.
 The velocity potential and the displacement of the linearised inviscid SW equations both satisfy the wave equation,
\begin{equation}
\frac{\partial ^2 \eta}{\partial t^2} = c^2 \nabla^2 \eta \, ,
\end{equation}
with equipartition between KE and APE over a wave period in each Fourier mode.

In order to analyze the spectral flux of energy, we derive the spectral energy
budget, i.e.\ the governing equations for spectral KE and APE
functions. This is not as straightforward as in incompressible turbulence since
the expression of the kinetic energy  is not
quadratic. To define the spectral KE function, we use the relation
\begin{equation}
\mean { E_K } = \sum_\kk \scalarprod{\uu}{\JJ}/2,
\end{equation}
where
\begin{equation}
\scalarprod{\baa}{\bb} \equiv \Re\{ \hat \baa(\kk)^* \cdot \hat \bb(\kk) \},
\end{equation}
and $\Re$ denotes the real part, $\kk$ is the wave number and the
hat denotes the Fourier transform.
%
The spectral KE function can therefore be defined as
\begin{equation}
E_K(\kk) \equiv \scalarprod{\uu}{\JJ}/2.
\end{equation}
Similarly, the APE can be written as the sum over all wave numbers of
the spectral APE function 
\begin{equation}
E_A(\kk) =  c^2 \scalarprod{\eta}{\eta}/2 \, .
\end{equation}
The inviscid equation for the spectral KE and APE functions can be written as
\begin{eqnarray}
\p_t E_K(\kk) &=& T_K(\kk) - C (\kk),\\
\p_t E_A(\kk) &=& T_A(\kk) + C (\kk),
\end{eqnarray}
where
\begin{equation}
T_K(\kk) 
= -\scalarprod{\bnabla \cdot \uu}{E_K} 
-\scalarprod{\uu }{\bnabla E_K } \, ,
\end{equation}
\begin{equation}
T_A(\kk) = -c^2 \scalarprod{\bnabla \eta }{\JJ}- c^2\scalarprod{\eta}{\bnabla \cdot \JJ}  \, ,
\end{equation}
are the spectral transfer functions of KE and APE, respectively, and
\begin{equation}
C (\kk) 
= \scalarprod{\uu}{\bnabla E_P}
\end{equation}
is the spectral KE-APE conversion function.
The spectral energy flux functions of KE and APE are defined as
\begin{eqnarray}
\Pi_K (k) = \sum_{| {\bf k}^{\prime} | > k} T_K ( {\bf k}^\prime ) \, , \;\;\;  \Pi_A (k) = \sum_{| {\bf k}^{\prime} | > k} T_A ( {\bf k}^\prime ) 
\end{eqnarray}

We will now derive an analogue of the so called `four-fifths  law' \cite[]{Kolmogorov1941} for the the third order velocity structure function of incompressible isotropic turbulence.   In the case of irrotational SW wave turbulence the analogous law contains a third order structure function involving both velocity and displacement increments.  In order for such a law to exist it is necessary that the velocity field is either non-divergent (incompressible) or irrotational. To derive the law it is convenient to start from the following equations
\begin{eqnarray} \label{IRR}
\p_t \uu &=& -\bnabla \cdot (\uu \cdot \uu /2) - c^2 \bnabla h  \\
\p_t \JJ &=& -\p_j (u_j \JJ) - \bnabla E_p,\\
\p_t h   &=& -\bnabla \cdot \JJ \, ,
\end{eqnarray}
where we have used the  assumption that the velocity field is irrotational in (\ref{IRR}). We consider two points, positioned at $ \bf x $ and $ {\bf x}^\prime $, separated by the vector $ {\bf r} = {\bf x}^\prime - {\bf x} $.
Using Cartesian tensor notation we denote derivates with respect to $ {\bf x} $ as $ \partial_i $ and derivatives with respect to $ {\bf x}^\prime $ as $ \partial^{\prime}_i $. Treating $ \bf x $ and $ {\bf x}^\prime $ as independent variables we find

\begin{equation}
\p_t ( J_i' u_i ) = - \p_i( J_i'|\uu|^2/2) - \p_i'(u_j J_j' u_i')
-\p_i (J_i' c^2 h)  - \p_i' (u_i E_p')
\end{equation}
and
\begin{equation}
\p_t ( h' h ) = - \p_i( h'J_i ) - \p_i'( hJ_i' ),
\end{equation}
which gives
\begin{eqnarray}
\p_t ( \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h )
&=& - \p_i( J_i'|\uu|^2/2) - \p_i'( J_i|\uu'|^2/2) %\nonumber \\&& 
- \p_i'(u_j J_j' u_i') - \p_i(u_j' J_j u_i) \nonumber\\
&& -c^2\p_i (J_i'h) - c^2\p_i' (J_ih') %\nonumber\\&& 
- \p_i' (u_i E_p') - \p_i (u_i' E_p) \nonumber\\
&& -2c^2\p_i' (J_i'h) - 2c^2\p_i (J_ih').
\end{eqnarray}
We then assume homogeneity and take the average denoted by
$\meane{}$.   For
every function $g(\xx, \xx')$, we have $\p_i' \meane{g} =
-\p_i\meane{g} = \bnabla_{\rr} \meane{g}|_i$. We thus obtain
\begin{eqnarray}
\p_t \meane{ \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h }
&=& + \bnabla_{\rr}\cdot\meane{ \JJ'|\uu|^2/2 - \JJ|\uu'|^2/2 } \nonumber\\
&& - \bnabla_{\rr}\cdot\meane{ u_j J_j' \uu' - u_j' J_j \uu } \nonumber\\
&& - \bnabla_{\rr}\cdot\meane{ \uu E_p'- \uu' E_p } \nonumber\\
&& -c^2\bnabla_{\rr}\cdot\meane{ \JJ'h - \JJ h' }
\end{eqnarray}
%
We then introduce the structure functions and the operator $\delta$
returning the increment of a variable between two points separated by $\rr$, 
for example $\delta h (\xx, \rr) = h(\xx+\rr)- h(\xx)$.
%
Using homogeneity again, we obtain
\begin{equation}
\meane{ (\delta h)^2 \delta\uu } = 
-\meane{ h'^2 \uu } + \meane{ h^2 \uu' }
+2\meane{ hh' \uu } - 2\meane{ hh' \uu' }
\end{equation}
and 
\begin{equation}
\meane{ |\delta \uu|^2 \delta\JJ } = 
\meane{  |\uu|^2 \JJ' } - \meane{ |\uu'|^2\JJ }
+\meane{ u_ju_j' \JJ' } - \meane{  u_ju_j' \JJ },
\end{equation}
which gives
\begin{equation}
2\p_t \meane{ \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h }
= \bnabla_{\rr} \cdot ( \meane{ |\delta \uu|^2 \delta\JJ } 
+ c^2\meane{ (\delta h)^2 \delta\uu } ).
\end{equation}
For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \cite[see for example][]{Frisch} gives that the left hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate. It is interesting to note that a factor of $ 8 $ appears in the case when the velocity field is irrotational while a corresponding factor of $ 4 $ appears in the case when the velocity field is non-divergent. 
Assuming isotropy and integrating yield the Kolmogorov law for
irrotational shallow water wave  turbulence
\begin{equation}
\meane{ |\delta \uu|^2 \delta J_L } 
+ c^2\meane{ (\delta h)^2 \delta u_L } = -4 \epsilon r, \label{eq_Kolmo}
\end{equation}
where $J_L \equiv
\JJ\cdot\rr / |\rr|$ and $u_L \equiv \uu\cdot\rr / |\rr|$ are
longitudinal increments.

We will now use (\ref{eq_Kolmo}) together with straightforward geometrical and statistical arguments to construct a model for velocity and displacement structure functions of any order above a certain minimum.
The model has some similarities with the one developed by \cite{BouchaudMezardParisi1995} for Burgers equation. The main difference is the that the mean dissipation rate, $\epsilon $, enters into the estimate of the shock amplitude and the final expressions of the structure functions in our model, but is absent in the model of \cite{BouchaudMezardParisi1995}.
We consider a domain with total area  $ {\mA} $ occupied by shocks which we 
consider as smooth lines. We denote the total length of all shocks by $ L_s $.  The mean distance between the shocks within the domain can be estimated as
\begin{equation}
d \simeq \frac{\mA} {L_s} .
\end{equation} 
A structure function of a flow variable is defined as the area average of increments  between two points separated by a distance $ r $.
Assuming that the contribution from increments where the two points are not separated by a shock is negligible for higher order structure functions we obtain
\begin{eqnarray}
\meane{|\delta h|^p} 
=  \frac{1}{\mA} \int_{\mA}  d^2\xx |\delta h(\xx)|^p 
 \simeq  \frac{r}{\mA} \int_\shocks ds |\sin\theta| |\Delta h (s)|^p,
\end{eqnarray}
where $\theta$ is the angle between the shock line and the separation
vector $\rr$ %
and $\Delta h (s)$ is the shock displacement amplitude.
%
Assuming isotropy, the integral can be split and we get
\begin{equation} \label{StepH}
\meane{|\delta h|^p} 
\simeq 
r \frac{L_s}{\mA} \meant{|\sin\theta|} \means{|\Delta h |^p}
\simeq  \frac{2}{\pi} \means{|\Delta h |^p} \frac{r}{d} ,
\end{equation}
where $\meant{}$ is the mean over $\theta$
and $\means{}$ is the mean over all shocks.

For velocity increments, we can also use a characteristic of the
velocity singularities.  The step in velocity is only in the component
perpendicular to the shocks, which implies that the longitudinal and
transverse increments can be written as  $\delta u_L = \delta u
\sin\theta$, $\delta u_T = \delta u \cos\theta$.
%
Applying the same averaging method, we obtain
\begin{equation} \label{StepL}
\meane{|\delta u_L|^p} 
\simeq 
\meant{|\sin\theta|^{1+p}}
\means{|\Delta u |^p} \frac{r}{d}
\end{equation}
and
\begin{equation} \label{StepT} 
\meane{|\delta u_T|^p} 
\simeq 
\meant{|\sin\theta||\cos\theta|^{p}}
\means{|\Delta u |^p} \frac{r}{d},
\end{equation}
where $\Delta u $ is the shock velocity amplitude.
%
These assumptions completely determine the ratio between the
longitudinal and transverse structure functions, which can be computed as
\begin{eqnarray} \label{Ratio}
R_p(r) \equiv \frac{\meane{|\delta u_L|^p}}{\meane{|\delta u_T|^p}} = 
\frac{\meant{|\sin\theta|^{1+p}}}{\meant{|\sin\theta||\cos\theta|^p}},
\end{eqnarray}
giving the numerical values $R_2 = 2$, $R_3 = 6\pi/8$ and $R_4 = 8/3$.
Interestingly, for an isotropic and irrotational velocity field, the longitudinal and transverse second order structure functions
are exactly related by $\meane{(\delta u_L)^2} = \p_r( r\meane{(\delta
u_T)^2} )$ \cite[]{Lindborg2007jas}. If we further assume than both second order structure functions follow
the same scaling law $\meane{ (\delta u_L)^2} \propto \meane{ (\delta
u_T)^2} \propto r^\alpha$, then we have $\meane{ (\delta u_L)^2} =
(1+\alpha) \meane{ (\delta u_T)^2} $, which gives $ R_2 = 2 $ for 
$  \alpha = 1$.  In the limit $ r \rightarrow 0 $ we have $ \alpha = 2 $ and $ R_2 = 3 $.
%

In order for (\ref{StepH}-\ref{StepT}) to match with the third order structure function law (\ref{eq_Kolmo})  the shock amplitudes should scale as
\begin{equation} \label{Strength}
 | \Delta u | \sim | c \Delta h | \sim (\epsilon d)^{1/3} \, ,
\end{equation} 
and the structure functions of order $ p $ will consequently scale as
\begin{equation} \label{StrucFunc}
\meane{|\delta u |^p}  \sim \meane{(c|\delta h |)^p} \sim  (\epsilon  d)^{p/3} \,  \frac{r}{d} \, .
\end{equation} 
Assuming that (\ref{StrucFunc}) is valid for second order moments and that the non-quadratic contribution to the KE spectrum is negligible the corresponding scaling law for the energy spectra reads
\begin{equation} \label{Spectra}
E_K(k)  \sim  E_A(k) \sim \epsilon ^{2/3} d^{-1/3} k^{-2} \, .
\end{equation} 
The  model  provides the following predictions for the flatness factors of velocity increments
\begin{eqnarray} \label{FL}
F_L &=&  \frac{\meane{|\delta u_L|^4}}{\meane{|\delta u_L|^2}^2}
\propto 
\frac{\meant{|\sin\theta|^5}}{\meant{|\sin\theta|^3}^2} \frac{d}{r}   = \frac{9\pi}{15}  \frac{d}{r} \\ \label{FT}
F_T &=&  \frac{\meane{|\delta u_T|^4}}{\meane{|\delta u_T|^2}^2} 
\propto
\frac{\meant{|\sin\theta||\cos\theta|^4}}{\meant{|\sin\theta||\cos\theta|^2}^2} =  \frac{9\pi}{10}  \frac{d}{r} 
\end{eqnarray}
The ratio between the two flatness factors is equal  to
$F_T/F_L = 3/2$ according to the model.

We now turn to the question of how the shock width, $ \delta x $, can be estimated. One suggestion would be that $ \delta x $  is equal to the Kolmogorov length scale, $ \nu^{3/4}/\epsilon^{1/4} $. However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
suggests that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $. The scaling of the energy spectra (\ref{Spectra}) also suggests that $ d $ should enter into the expression for $ \delta x$, apart from $ \nu $ and $ \epsilon $. Dimensional analysis thus gives
\begin{equation} \label{Width}
\delta x \sim \frac{\nu}{\; (d\epsilon)^{1/3}} \, ,
\end{equation} 
which is considerably smaller than the Kolmogorov length scale.
This result can also be derived in the following way. Assuming that all dissipation takes place at the shocks, whose total area is equal to $ L_s \delta x $,  and remembering that $ \epsilon $ is defined as the mean over the domain area $ \cal  A$, we find
\begin{equation} \label{Estimate}
\epsilon \sim \nu \frac{ \mid \Delta u \mid ^{2} }{\delta x^2} \frac{L_s \delta x}{\cal{A}} \sim \nu \frac{\epsilon^{2/3}} {\delta x d^{1/3}} \, ,
\end{equation} 
again giving (\ref{Width}). 
Applying the same averaging method we obtain the following estimates for skewness and flatness of the velocity derivative, $ \partial_x u $,
\begin{eqnarray}
S_{\partial_x u}  & \sim & \left ( \frac{d}{\delta x} \right )^{1/2} \sim \frac{d^{2/3} \epsilon ^{1/6}}{\nu^{1/2}} \, , \\
\label{FD} F_{\partial_x u} & \sim &  \frac{d}{\delta x} \sim \frac{d^{4/3} \epsilon^{1/3}} {\nu} \, ,
\end{eqnarray} 
where the skewness is negative. For Kolmgogorov turbulence observations indicate that $ F_{\partial_x u} \propto \nu^{-1/6} $  \cite[]{VanAtta_Antonia1980}, a dependence which is well capture by the multifractal model of
\cite{Meneveau_Streenivasan1991}.  In SW wave turbulence there is a much stronger increase of  the flatness factor with decreasing  $ \nu $, which can be explained the fact that dissipation takes place only at the shocks.

Finally, we estimate the Mach number of the shocks, that is the ratio between the shock propagation velocity and $ c $. According to \cite{Baines1998} the Mach number of a SW shock can be calculated as
\begin{equation}
M = \sqrt{\frac{h_+}{h_-} \frac{h_+  + h_-}{2}} \, ,
\end{equation} 
where $ h_+ $ and $ h_- $ are the nondimensional layer thickness before and after the shock, respectively. Noting that $ h_+ - h_- =  \;  \mid \!\! \Delta h | $, expanding in small $ | \Delta h | $ and using (\ref{Strength}) the Mach number can be estimated as
\begin{equation} \label{Mach}
M \sim 1 + \frac{3}{4} \frac{\;\; (\epsilon d)^{1/3}}{c} \, ,
\end{equation} 
and is only slightly larger than unity for all relevant applications.

The model we have developed is, indeed, very simple. As such, it must, of course, have limitations. First, it is obvious that  the linear scaling of the structure functions only can be valid for separations  that fulfil the condition $ \delta x \ll r \ll d $. Second, the model cannot be valid for low order moments or structure functions. The minimum order for which the model may be valid is not easily determined. 
According to  \cite{BouchaudMezardParisi1995} structure functions of shock dominated turbulence of all orders $ p  \ge 1 $ scale linearly in $ r $. It is quite easily argued, however, that the shocks can not determine the scaling of the structure functions of order $ p=1 $. 
To see this, consider the first order moment of the absolute value of $ \delta u_L $. 
The longitudinal velocity increment is always negative over a shock, that is to say that the difference of the projection of the velocity on the direction over which the shock is crossed is always negative. In a domain with periodic boundary conditions the mean value of $ \delta u_L $   is zero. The negative values of $ \delta u_L $ for separations where a shock is crossed must therefore be compensated by positive values for separations where no shock is crossed. There is no reason to believe that these values are all positive,  but they are only positive on average. Thus, 
$ \means{ |\delta u_L | } < \langle | \delta u_L | \rangle _{ns} $, where the left hand side is the mean value over all separations where a shock is crossed and the right hand side is the mean value over separations where no shock is crossed.  Therefore, the model is not valid for first order moments. It may be conjectured that it is valid for 
second and higher order structure functions.

\section{Simulations}

We have carried out a series of twenty forced-dissipative simulations of the SW
equations in a quadratic domain with periodic boundary conditions using the
standard open-source pseudo-spectral code FluidSim \cite[]{fluiddyn, fluidfft,
fluidsim}.
%
The simulations were started from zero initial
conditions and were run till a statistically stationary state was reached. The
equations were reformulated in terms of the three eigenmodes of the linearized
equations in Fourier space. One of these eigenmodes is zero for an irrotational
flow. We first carried out a few test simulations in which all three eigenmodes
were allowed to evolve in time. However, when only the two wave modes are
forced vorticity remains very small throughout the simulations and does not
affect the wave dynamics. Therefore, we chose to solve the equations only for
the two wave modes. As discussed in the previous section, a linear viscous term
in (\ref{eq_uu}) will not conserve momentum exactly and does not give a
strictly positive definite expression for energy dissipation. We have relaxed
the requirements of exact momentum conservation and positive definite energy
dissipation. Instead of using the nonlinear form of the viscous term in
(\ref{eq_uu}) we have used a higher order linear operator $ \nu_8 \nabla^{8} $
both in (\ref{eq_uu}) and (\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion'
or a `hyperviscosity'. There is no strong physical motivation for this
modification but, as discussed by \cite{FargeSadourny1989}, a molecular viscous
operator of the form $\nu \nabla^2 $ is not necessarily a relevant model of the
actual small-scale dissipation for SW flows, which should rather be described
as a transition from two-dimensional to three-dimensional motions before
reaching the scales where dissipation actually occurs.
%
Time advancement is carried out by a classical fourth-order Runge-Kutta scheme
for the nonlinear term and an exact integration for the linear and dissipative
terms. This explicit integration is especially useful at very high resolution
and very large $c$ when the shortest waves are very fast. We use an adaptable
time step method which maximizes the time step over a standard
Courant-Friedrichs-Lewy condition \cite[]{Lundbladh1999,
AugierChomazBillant2012}.
%
Most of the aliasing is removed by truncating 8/9 of the modes along each
direction \cite[for a detail discussion on the issues of the non-conservation
of the non-quadratic energy and the aliasing errors in the truncated one-layer
shallow water model, see][]{FargeSadourny1989}. The resolution is characterized
by the number of nodes, $n$, in each direction and has been varied from 960
to 7680.
%
The wave speed $c$ has been varied over almost two orders of magnitude from 10
to 700. The wave modes are forced in a shell in spectral space corresponding to
relatively small wave numbers, $ 5 \delta k < |\kk| < 8 \delta k$, where $
\delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic
domain. 
The forcing is white noise in time and designed so that the energy
injection rate is normalised to unity at each time step. When the flow has reached a stationary state dissipation should therefore be equal to unity. 
The strength of the stratification is characterized by a
Froude number, defined as
\begin{equation} \label{Fr}
F_f = \frac{\epsilon^{1/3}}{c k_f^{1/3}} \, ,
\end{equation}
a parameter that can be determined from the start of each simulations.
Energy is dissipated at the smallest resolved scales by hyperviscosity. An intricate question is how the value of $ \nu_{8} $ should be chosen so that the shocks are resolved. The standard resolution requirement in turbulence simulations of this type is that the hyperviscosity analogue to the Kolmogorov scale
\begin{equation}  \label{KolmEight}
\eta_{8} = \left( \frac{{\nu_8}^3}{\epsilon} \right)^{\frac{1}{22}},
\end{equation}
should be resolved. However, our estimate (\ref{Width}) of the shock width suggests that this may not be a sufficient condition, since $ d $ is likely to enter the expression for $ \delta x $ also in the case when we use hyperviscosity.
A similar estimate as in (\ref{Estimate}) gives
\begin{equation} \label{EtaEight}
\delta x \sim \eta_{8} \left ( \frac{\eta_8}{d} \right)^{1/21} \, .
\end{equation} 
Thus, the shock width dependence on $ d $ can be expected to be extremely weak when hyperviscosity is used. Therefore we simply use the standard resolution requirement that (\ref{KolmEight}) should be resolved. We have found that setting
$  k_{d} \simeq \kmax/ 2.5$, where $ k_{d} = 1/\eta_8 $ and $\kmax =
(8/9)\pi n/L $ is the maximum wave number, is  sufficient to resolve the shocks. Since the energy injection rate is normalised to unity and all simulations reach a stationary state in which dissipation is equal to the injection rate, $ \nu_8 $ can be chosen already from the start so that this condition will be fulfilled in the stationary state. 
 The  simulation parameters are listed in table 1.

\begin{table}
\begin{center}

\label{Table1}
\begin{tabular}{lrrrrrrrr}

\toprule
{} &   $n$ &  $c$ &  $\nu_8$ &  $\epsilon$ &  $ k_{d}  / {k_f} $ &   $F_f$ &    $t_{\max}$ \\
\midrule
W1  &   960 &   10 & 1.56e-10 &        1.04 &                    28.8 &   0.111 &                49.9 \\
W2  &  1920 &   10 & 9.68e-13 &        1.06 &                    57.7 &   0.112 &               49.8 \\
W3  &  3840 &   10 &    6e-15 &         1.10&                     115 &   0.113 &               49.9 \\
W4  &  7680 &   10 & 3.72e-17 &        1.18 &                     231 &   0.116 &               49.9 \\
W5  &   960 &   20 & 1.56e-10 &           1.00 &                    28.8 &   0.055 &              49.9 \\
W6  &  1920 &   20 & 9.69e-13 &        1.02 &                    57.7 &  0.0553 &                120 \\
W7  &  3840 &   20 & 6.01e-15 &        1.08 &                     115 &  0.0563 &                120 \\
W8  &  7680 &   20 & 3.72e-17 &        1.21 &                     231 &  0.0585 &                 120 \\
W9  &   960 &   40 & 1.56e-10 &          1.00 &                    28.8 &  0.0274 &                49.8 \\
W10 &  1920 &   40 & 9.68e-13 &       1.00 &                    57.7 &  0.0274 &               49.8 \\
W11 &  3840 &   40 &    6e-15 &        1.06 &                     115 &   0.028 &               49.8 \\
W12 &   960 &  100 & 1.56e-10 &       1.00 &                    28.8 &   0.011 &                49.8 \\
W13 &  1920 &  100 & 9.69e-13 &        1.01 &                    57.7 &   0.011 &                 120 \\
W14 &  3840 &  100 & 6.01e-15 &        1.06 &                     115 &  0.0112 &                 120 \\
W15 &  7680 &  100 & 3.72e-17 &        1.22 &                     231 &  0.0117 &                95.8 \\
W16 &   960 &  400 & 1.56e-10 &        1.14 &                    28.8 & 0.00287 &                 120 \\
W17 &  1920 &  400 & 9.69e-13 &        1.15 &                    57.7 & 0.00288 &                 120 \\
W18 &  3840 &  400 & 6.01e-15 &        1.12 &                     115 & 0.00286 &                119 \\
W19 &   960 &  700 & 1.56e-10 &        1.27 &                    28.8 &  0.0017 &                49.8 \\
W20 &  1920 &  700 & 9.68e-13 &        1.05 &                    57.7 &  0.0016 &                49.8 \\
\bottomrule

\end{tabular}
\caption{Parameters for all simulations. $ n $: number of nodes in each
direction, $ c $: wave speed, $ \nu_8 $: hyperviscosity, $ \epsilon $: time
averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio
between dissipation wave number and forcing wave number, $ F_f $: Froude
number, $ t_{max} $: end time of simulation.}
\end{center}
\end{table}

\begin{figure}
\centerline{\includegraphics[width=5.12in]{../Pyfig/fig_Emean_time}}
\caption{Space averaged energy, $E = E_K + E_A $, versus time for all
simulations. }
\label{fig_Evstime}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=5.8in]{../Pyfig/fig_energy_w}}
\caption{Left: Mean energy in the stationary state versus $ c $ for different
resolutions. The lines have slope $ 0.5 $, corresponding to $ E \propto c^{1/2}
$. Right: Normalised mean energy in the stationary state versus resolution $ n
$ for different $ c $.}
\label{MeanE}
\end{figure}

The time evolution of mean energy is shown in figure~\ref{fig_Evstime} for all
twenty simulations. After an initial period of growth, mean energy is leveling
out at different values for different runs. This is not the same picture as we
see in simulations of three-dimensional incompressible turbulence or stratified
turbulence, where all curves would have leveled out more or less at the same
value. We find that mean energy in the stationary state is a function of both $
c $ (or the Froude number) and resolution $ n $ and that the functional
dependence can be approximately written as
\begin{equation} \label{MeanEnergy}
E \sim \sqrt{L_f c \epsilon} \, n^{\alpha}  \, ,
\end{equation}
where $ L_f = \pi/k_f $ is the forcing scale. In figure~\ref{MeanE} to the
left, we see $ E/\sqrt{\epsilon} $ as function of $ c $ for different resolutions $
n $. For each resolution the points are approaching a power law $ C_n c^{1/2} $
with increasing values of $ C_n $ for increasing $ n $. In the same figure to
the right we see $ E/\sqrt{L_f c \epsilon} $ for different values of $ c $. For
each $ c $ the points follow a power law $ n^{\alpha} $, where $ \alpha $ is
becoming slightly larger with increasing $ c $, approaching almost the same
value $ \alpha \approx 0.33 $ for the three highest values of $ c $. The
relation (\ref{MeanEnergy}) can be reformulated as
\begin{equation} \label{Dissipation}
\frac{\epsilon} {(E^{3/2} / L_f)} \sim \frac{E^{1/2}}{c} n^{-2\alpha} \, .
\end{equation}
Since the simulations are well resolved the dependence on resolution should not
be interpreted as consequence of the numerics but rather as a Reynolds number
effect, although it is not hundred percent clear how a Reynolds number should
be defined in our case. We may note, however, that $ n \sim L_f/\delta x $ and in the case of Navier-Stokes viscosity
 we should have $ \delta x \propto \nu $, according to (\ref{Width}). Therefore,  $ n $ may be substituted by a Reynolds number. The normalised energy dissipation on the left hand side of
(\ref{Dissipation}) will thus go to zero in the limit of large Reynolds number.
This is very different from three-dimensional turbulence where there is a local
Richardson-Kolmogorov cascade, in which energy is successively transfered from
large to small scales of motion. That the cascade is local means that eddies of
a particular scale are not directly influenced by eddies of widely different
scales. As a consequence, the normalised dissipation is of the order of unity
independent of Reynolds number \cite[]{Pope, TennekesLumley}. A review of the
experimental evidence and theoretical implications of this fundamental law of
Kolmogorov turbulence is given by \cite{Vassilicos2015}. Our result
(\ref{Dissipation}) suggests that SW wave turbulence does not fit into the
paradigm of a local Richardson-Kolmogorov cascade. Moreover, it is not only in
the limit of large Reynolds number the normalised dissipation will go to zero
but also in the limit of strong stratification, $ E^{1/2}/c \rightarrow 0 $.
This is different from three-dimensional stratified turbulence where the ratio
on the left hand side of (\ref{Dissipation}) is of the order of unity in the
limit of strong stratification \cite[]{Lindborg2006, Brethouwer2007}.

The `four-fifths law' for the third order structure function of Kolmogorov
turbulence is often derived using the assumption of a local cascade
\cite[]{Vassilicos2015} or the assumption that dissipation stays finite in the
limit of large Reynolds number \cite[]{Frisch}. As we just discussed, these
assumptions seem to be violated in the case of SW wave turbulence.
Nevertheless, the analogous law (\ref{eq_Kolmo}) is satisfied with a high degree of accuracy in our simulations. In
figure~\ref{Flux} we see the spectral energy flux from run W7 to the left and
the third order structure function to the right. There is a broad constant flux
range with equipartition between KE and APE flux and an almost equally broad
range where (\ref{eq_Kolmo}) is satisfied, with equipartition between the two
third order structure functions associated with KE and APE. The assumptions of
a local cascade and finite dissipation do not seem to be necessary in order for
the third order structure function law (\ref{eq_Kolmo}) to be valid.

\begin{figure}
\centerline{\includegraphics[width=5.8in]{../Pyfig/fig_flux_struct_combined}}
\caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f
$. Right: Time averaged normalised third order structure functions versus $
r/L_f $. From run W7. }
\label{Flux}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=3.15in]{../Pyfig/fig_spatiotempspectra}}
\caption{Frequency spectra of KE and APE for different $ k = \mid {\bf k} \mid
$, from run W7. The spectra are plotted as functions of $\omega/\omega_l$,
where $\omega_l = c k$. From top to bottom: $ k /\delta k = 12$, 27, 62, 143,
327, and 746. }
\label{fig_spatiotemp_spectra}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=6.0in]{../Pyfig/fig_phys_fields_wave}}
\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
\label{Physical}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=8cm]{../Pyfig/fig_shock_sep.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance}
\end{figure}

The equipartition of KE and APE spectral fluxes suggests that the flow to
leading order may be described as a collection of linear gravity waves with energy
equipartition in each mode. To investigate this further we have
collected time series of the flow variables from Fourier modes whose magnitude
fall within certain shells, $ \mid {\bf k} \mid \in [ k -\delta k/2, \; k+
\delta k/2] $, and computed temporal power law spectra from these time series
and averaged over all wave numbers in each shell. In other words, we have
computed KE and APE frequency spectra corresponding to different magnitudes of
modes. Figure~\ref{fig_spatiotemp_spectra} presents such spectra plotted as 
functions of the normalised frequency $\omega/\omega_l$, where $ \omega_l = kc
$ is the linear wave frequency. The spectra are strongly dominated by peaks at
$\omega = \omega_l$. As can be seen, there is equipartition between KE and APE
in each shell. Although the widening of the peaks around the linear frequency
most likely is an effect of nonlinearities, we can quite safely conclude that
the time evolution of the flow variables in each Fourier mode to leading order
can be described as a linear wave.

Visualisations of the flow field give a completely different picture from what
may be expected from a collection of linear gravity waves. They are totally
dominated by the appearance of shock waves. In figure~\ref{Physical} we see
the divergence, $ \nabla \cdot {\bf u} $, to the left and the velocity
component in the $ y $-direction, $ u_y $, to the right. The two figures at the
top are from run W6 while the two figures at the bottom are from run W17. The
difference between the two runs is that $ c $ is larger by a factor of $ 20 $
in W17 as compared to W6. To the left we see the shocks displayed as elongated
bands of negative divergence. That the divergence is negative at the shocks can
be understood from the fact that the velocity component perpendicular to the
shock always has a negative jump in the direction of shock propagation.
Apparently, the mean distance between the shocks is considerably smaller in run
W17 with the larger value of $ c $ compared to run $ W6 $ with the smaller
value of $ c $. 
The shocks are also visible in the figures to right where $ u_y
$ is displayed. It may be interesting to note that the shocks along the $ x
$-axis appear as extra sharp in this plot, since there is a strong jump of $
u_{y} $ over these shocks, while the shocks along the $ y $-axis are hardly
visible, since there is no jump of $ u_{y} $ over these. In the two figures to
the right we see a variation of the flow field over the length scale $ L_f $
which is not seen in the two figures to the left. This variation is evidently a
footprint of the random forcing. Although the forcing length scale $ L_f $ is
the same in the two simulations the mean distance, $ d $, between the shocks is
much smaller in run $ W17 $ than in run $ W6 $, which may come as a surprise. Another interesting observation is that the shocks in run W17 (lower figures) appear to have smaller curvature on average than the shocks in run W6 (upper figures). In both runs, the shocks are weak in the sense that the Mach number is close to unity and the interaction between the shocks should therefore be weak \cite[see][]{ApazidisEliasson2018}. However, as measured by the Mach number  (\ref{Mach}), the shocks are much weaker in run W17, where $ c $ is larger by a factor of 20 as compared to run W6 and $ d $ also is smaller. The shock interaction should therefore be weaker in run W17. The observation that the curvature of the shocks is smaller in run W17 than in run W6 may be explained by this fact. 
We have calculated  the mean distance, $ d $, between the shocks by counting the number of negative spikes
of the divergence along ten lines parallel to the $ x $-axis and ten lines
parallel to the $ y $-axis and then divided total length of all the lines by
the number of spikes. Generally, we found that $ d \propto c^{-1/2} $, or
\begin{equation} \label{MeanDistance}
d \sim L_f F_f ^{1/2} \, ,
\end{equation}
independent of resolution $ n $. In figure~\ref{fig_distance} we have plotted
$ d/L_f $ versus Froude number for all twenty runs. As can be seen, the points
follow the $ F_f^{1/2} $-curve quite well, although there seems to be a slight
systematic deviation at the highest $ F_f $, which may be due to the fact that
$ d $ is rather close to the length of the side of the box for the largest $
F_f $.

\begin{figure}
\centerline{
\includegraphics[width=5.0in]{../Pyfig/fig_struct_order_246}}
\caption{From top to bottom: second, fourth and sixth order compensated and
normalised longitudinal structure functions. Left: runs with $ n=3840 $: W3,
W7, W11 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15. }
\label{fig_StrucFunc}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=11cm]{../Pyfig/fig_ratio_strfct}}
\caption{
Ratio of the structure functions
$R_p(r) = \mean{|\delta u_L|^p} / \mean{|\delta u_T|^p}$ from run W4,
for $ p = 2, \, 3, \, 4, \, 5, \, 6. $
The dotted straight lines indicate the values predicted by (\ref{Ratio}):
$R_2 = 2$, $R_3 = 3\pi/4$,  $R_4 = 8/3$, $ R_5 = 15 \pi /16 $ and $ R_6 = 16/5 $.}
\label{fig_ratio}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=12cm]{../Pyfig/fig_flatness}}
\caption{ Top: Flatness factor of transverse velocity increments. Bottom: Ratio between transverse and longitudinal flatness factor. Left: runs with $ n=3840 $: W3,
W7, W11 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15.  The dotted straight lines indicate the value 1.5 predicted by the shock model.  }
\label{fig_flatness}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=11cm]{../Pyfig/fig_spectra}}
\caption{Time averaged compensated and normalised energy spectra
from runs with $ n = 7680 $: W4, W8 and W15.}
\label{fig_spectra_c40}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=13cm]{../Pyfig/fig_diss_spectra}}
\caption{Time averaged normalised dissipation spectra. Left: Constant $ c = 20 $ and varying $ n $, runs  W5, W6, W7 and W8. Right: Constant $ n= 1920 $ and varying $ c $, runs W2, W6, W10, W13, W17 and W20. The vertical lines indicate the wave numbers of the maxima.}
\label{fig_disspectra}
\end{figure}

We will now present the results for the structure functions. Substituting
(\ref{MeanDistance}) into (\ref{StrucFunc}) we find that the structure
functions are expected to scale as
\begin{equation} \label{StrucFunc2}
\meane{|\delta u |^p}  \sim \meane{(c|\delta h |)^p} \sim  (L_f F_f^{1/2})^{p/3-1}  \epsilon^{p/3}  r  \, .
\end{equation}
for separation distances that fulfil the condition $ \delta x \ll r \ll d $.
The scaling (\ref{StrucFunc2}) is
expected to become better and better fulfilled with decreasing $ c $ (or
increasing $ F_f $). According to (\ref{MeanDistance})
$ d $ is increasing with $ F_f $, and the scaling range is therefore becoming
wider for decreasing $ c $, if $ \delta x $ or the resolution  is kept
constant. The scaling
(\ref{StrucFunc2}) is also expected to become better and better fulfilled with
increased resolution $ n $, since $ \delta x $ is decreasing with increasing $
n $.

As expected, we have found that the displacement structure functions scale
exactly in the same way as the velocity structure functions, why we only plot
velocity structure functions. In figure~\ref{fig_StrucFunc} we see the
compensated and normalised second, fourth and sixth order longitudinal
structure functions from the five simulations with $ n = 3840 $ to the left and
the three simulations with $ n = 7860 $ to the right. At small separation, $ r
< \delta x $, the structure functions have the dependence $ \sim r^{p} $, which
is characteristic of a smooth velocity field, indicating that the shocks are
reasonably well resolved. There is a small bump around the shock width, $ r
\approx \delta x $, and at $ r > \delta x $, there is a range where the
compensated structure functions are flat, or almost flat. As expected, this
range is becoming broader for smaller $ c $ (or larger $ F_f $). The range is
also broader in the higher resolution runs to the right as compared to the
lower resolution simulations to the left. There is clearly a better collapse of
the curves for the sixth order structure functions compared to the second order
structure functions. This is not surprising, since the shock contribution to
the structure functions should become increasingly dominant with increasing
order. Plots of the transverse structure functions are very similar to the
plots of the longitudinal structure functions, with the difference that the
transverse structure functions are somewhat smoother at $ r \approx \delta x $.
In figure~\ref{fig_ratio} we see the ratio $ R_{p}(r) = \langle \mid \delta
u_L \mid ^{p} \rangle / \langle \mid \delta u_T \mid ^{p} \rangle$ for $ p=
2,3,4,5,6 $ from run W4, together with dotted straight lines, indicating the
values predicted by (\ref{Ratio}). As can be seen, there is a quite good
agreement with the predicted values in a limited range of separations, $ r >
\delta x $. Somewhat unexpectedly, the range in which there is a good agreement
becomes more narrow with increasing $ p $, an observation that we cannot
explain. It may also be noted that $ R_2 \rightarrow 3 $ in the limit of small
$ r $, which is the theoretical single point limit for an irrotational
isotropic field. In figure~\ref{fig_flatness} we se the flatness factor, $
F_T $, of the transverse structure functions at the top and the ratio, $
F_T/F_L $, at the bottom. The figures to the left show the results from the
five simulations with $ n = 3840 $ and the figures to the right show the
results from the three simulations with $ n = 7860 $. The flatness factor shows
increasingly large values for decreasing $ r $ with a dependence that is rather
close to but somewhat steeper than $ r^{-1} $. At small $ r $ the curves are
leveling out at very large values. For $ n = 7869 $ and $ c = 10 $ we see that
$ F_{T} \approx 400 $ at small $ r $. The degree to which the flatness factor
is increasing with decreasing scale is by some authors defined as the most
appropriate measure of spatial intermittency \cite[see for example][]{Frisch}.
If this measure is used, SW wave turbulence is, indeed, extremely intermittent.
In the two figures at the bottom we see that $ F_{T}/F_{L} \approx 1.5 $ in a
limited range of separations, $ r > \delta x $, which is becoming somewhat
broader as $ c $ is increasing and is also somewhat broader in the higher
resolution runs shown to the right as compared to the lower resolution runs
shown to the left. This is in accordance with the shock model predictions
(\ref{FL}) and (\ref{FT}).

Finally, we present energy and dissipation spectra. As expected, we found that APE and KE
spectra were almost identical, why we only plot the
total spectra. In figure~\ref{fig_spectra_c40} we see  compensated and
normalised energy spectra from the three highest resolution runs. The
compensated spectra are almost flat in a broad range. In the high wave number
end the compensated spectra show a `bottleneck', which we interpret as a
signature of the shocks. The $ k^{-2} $-dependence and the associated scaling predicted
by (\ref{Spectra}) are very well satisfied in a limited range of high wave
numbers close to the bottleneck. At smaller wave numbers the spectra are
becoming somewhat steeper than $ k^{-2} $ and the spectrum from the run with
largest $ c $ does not collapse perfectly onto the other two spectra.  In figure \ref{fig_disspectra}
 we see normalised dissipation spectra for fixed $ c= 20 $ and different $ n $ to the left, and fixed $ n = 1920 $
and different $ c $ to the right. First we note that the spectra go to zero quite nicely for large $ k $, indicating that the shock width $ \delta x $ is well resolved. Second, we note that there is a slight shift of the maximum in both figures, even though we have normalised the wave number by $ k_d = 1/\eta_8 $.   These shifts may be explained by the estimate (\ref{EtaEight}) predicting that $ \delta x $ should vary slightly with $ d $. If the maximum of the dissipation spectra is associated with $ \delta x $, there should be a shift of the maximum  to the right with increasing $ n $ and to the left with increasing $ c $, since $ d \propto c^{-1/2} $.
 The ratio between the maximum and the minimum $ n $ in the left figure is equal to 8 and  according to (\ref{EtaEight}) the corresponding ratio between the wave numbers at which the maxima are observed should be $ 8^{1/21} \approx 1.1 $, which is exactly the value we observe. The ratio between the maximum and minimum $ c $ in the right figure is equal to $ 70 $. Assuming that $ d \propto c^{-1/2} $ the ratio between the wave numbers at which the maxima are observed should be $ 70^{1/42} \approx 1.1 $, while the observed ratio is equal to $ 1.16 $.

\section{Conclusions}

We have carried out a theoretical and numerical investigation of SW wave
turbulence. First, we derived the SW analogue (\ref{eq_Kolmo}) of the
`four-fifths' law of Kolmogorov turbulence. Using this relation and
straightforward statistical and geometrical arguments we developed a simple
shock model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3}
$, where $ d $ is the mean distance between the shocks, and that the $ p $:th
order structure function above a certain minimum will scale as $ (\epsilon
d)^{p/3} r/d $. Then, we carried out a series of twenty forced dissipative
simulations, varying the Froude number and the resolution. In all simulations,
a statistically stationary state was reached and in all simulations we
observed shocks. The flow variables in each Fourier mode were found to evolve
in accordance with linear wave dynamics, with equipartition between APE and KE
over a period. The third order structure function relation was fulfilled with a
high degree of accuracy. From the simulations we made two interesting
observations that fall outside the predictive scope of the model. The first
observation was that mean energy in the stationary state approximately scales
as (\ref{MeanEnergy}), suggesting that the normalised dissipation
(\ref{Dissipation}) will go to zero both in the limit of small viscosity and
large wave speed. This is different from three-dimensional Kolmogorov
turbulence as well as strongly stratified turbulence, where, in both cases,
dissipation is finite in the limit of small viscosity, and in the latter case
also is independent of the degree of stratification. This observation suggests
that SW wave turbulence does not fit into the paradigm of a local
Richardson-Kolmogorov cascade. The second observation we made was that the mean
distance between the shocks scales as $ c^{-1/2} $, if the other parameters are
held constant. As we increased $ c $, the shocks were thus becoming denser.
Using this observation we tested the shock model by plotting the normalised and
compensated structure functions of different orders. Generally we found that
there was a quite good agreement between the simulation results and the model
predictions, becoming better with decreasing wave speed and increasing
resolution.

Combining the model and the observations we can get a quite good picture of the
dynamics in the limit of weak nonlinearities, $ F_{f} \rightarrow 0 $. The
picture may become more vivid if we say `wave breaking' were we previously have
talked about shock formation, and also say `wave lengths' were we previously
have talked about length scales. We identify the longest wave length, $
\lambda_l $, with the forcing scale $ L_f $ and the breaking wave length, $
\lambda_b $, with the mean distance, $ d $, between the shocks. There are two
dynamically different regimes, the short wave regime, $ [\delta x, \;
\lambda_b] $, and the long wave regime, $ [\lambda_b, \, \lambda_l] $. We first
consider the latter, whose width scales $ \lambda_{l} /\lambda_b \sim
F_f^{-1/2} $. Identifying the amplitude of the breaking waves with the shock
amplitude, $ (\epsilon d)^{1/3} $, we find that the ratio between the amplitude of
the longest waves and the breaking waves scales as $ E^{1/2}/(\epsilon d)^{1/3}
\sim F_{f}^{-5/12} n^{\alpha/2} $, where we may substitute $ n $ with a
Reynolds number. Thus, for small $ F_f $ and large Reynolds
number the longest waves are not directly influenced by wave breaking. The
Reynolds number dependence of the amplitude ratio makes it unlikely that a
general similarity theory can be easily formulated for this range. Therefore,
predictions from weak turbulence theory \cite[for
example][]{ZakharovSagdeev1970} are not likely to hold for this range, although
it is not directly influenced by wave breaking. The exact values of the power
law exponents in the argument leading to this conclusion are, of course, not
important. As for the short wave regime, the width scales as $ \lambda_b
/\delta x \sim F_f^{1/2} $ for fixed $ \delta x $ or fixed Reynolds number. As
$ F_{f} \rightarrow 0 $ the range will eventually become so narrow so that
breaking will disappear. Clearly, a necessary condition for wave breaking is 
\begin{equation} \label{Condition} 
\frac{\lambda_b}{\delta x} \gg 1 \, .
\end{equation} 
It is interesting to note that this condition means that the flatness factor of velocity derivatives (\ref{FD}) should be much larger than unity and that this flatness factor is equal to the longitudinal flatness factor (\ref{FL}) in the limit $ r \rightarrow 0 $.
Substituting (\ref{Width}) into (\ref{Condition})  and assuming that (\ref{MeanDistance}) is valid also in the case when Navier-Stokes viscosity is used we can formulate
the condition as
\begin{equation} 
Re F_f^{2/3} \gg 1 \, ,
\end{equation} 
where we have defined the Reynolds number as $ Re  \equiv \epsilon^{1/3} L_{f}^{4/3} / \nu $. It may be conjectured  that such a condition is sufficient for wave breaking. If this is true, 
wave breaking will appear for any small $ F_f $ provided that the Reynolds number is sufficiently large.

The question that originally motivated this study was if SW wave turbulence can
be used as a model for energy dissipation in the atmosphere and the oceans.
Undoubtedly, the answer to this question is negative and the reason for this is
not primarily the appearance of shocks and the associated $ k^{-2} $-spectrum
which is not consistent with the observed $ k^{-5/3} $-spectrum. The main
reason is that SW wave turbulence is a far too inefficient agent of
dissipation. Since the Reynolds number of geophysical flows is extremely large
and the stratification is strong, the relation (\ref{Dissipation}) implies that
dissipation would be absent in the interior of the atmosphere and the oceans
with SW turbulence as the responsible agent. This not what we observe. In a companion paper \cite[]{LindborgMohanan}
we develop a toy model of geophysical turbulence which is similar to the SW equations in the mathematical structure, but is absent of shocks. 
In this model energy is dissipated through a local cascade involving both vortical and wave modes.

Although the answer to the question that originally motivated this study is
negative, it may lead us or some other investigators to pursue another
interesting research direction. As pointed out in the introduction, the
acoustic equations are to the lowest order expansion in density perturbations
identical to the SW equations. This study may therefore give important clues to
the long standing problem of acoustic turbulence, in two and three dimensions.
A similar shock model as we developed in two dimensions may easily be
formulated in three dimensions, where the third order structure function law
will be similar to (\ref{eq_Kolmo}), with the only difference that the constant
$ 4 $ on the right hand side should be replaced by $ 8/3 $. In three dimensions
we find
\begin{equation}
\meane{ |\delta \uu|^2 \delta J_L }
+ c^2\meane{ (\delta h)^2 \delta u_L } = - \frac{8}{3} \epsilon r. \label{eq_Kolmo3}
\end{equation}
In the case of acoustic turbulence $ \delta h $ should be interpreted as a
normalised density increment rather than a displacement increment. If shocks in
three dimensions are assumed to form as smooth surfaces rather than smooth
lines, then the mean distance between the shocks will appear in the
model as $ d \sim {\cal{V}}/A_s $, where $ A_{s} $ is the total shock area and
$ \cal{V} $ is the volume of the domain. The shock amplitude will thus scale as
$ (\epsilon d)^{1/3} $ and the structure functions will scale as $ (\epsilon
d)^{p/3} r/d $ also in three dimensions, provided, of course, that shocks will
form in a similar way as in two dimensions. Generally, shock waves have similar 
stability properties in two and three dimensions \cite[]{ApazidisEliasson2018, LivertsApazidis2016} and 
it can therefore be expected that the three-dimensional case should be similar to the two-dimensional case. 
If it is, the most interesting
questions are if the mean energy and the mean distance between the shocks scale
in a similar way in three dimensions as in two dimensions. Simulations or
experiments of acoustic turbulence may provide answers to these questions --
answers that will lead to further clues to a theoretical understanding of
acoustic turbulence. We hope that the present study will stimulate research in
this direction.

%\appendix

%\section{Kinetic energy dissipation and forcing rates}
%\label{app_comp}
%\input{appendix_dissip_forcing}

\bibliographystyle{jfm}
\bibliography{biblio}

\end{document}